---
title: Integration of omics and clinical data of the hormonal, metabolic, inflammatory,
  and oxidative response in the different macronutrients of the diet
author: "Edmond Geraud"
subtitle: '`r params$subtitulo`'
date: '`r format(Sys.Date(),"%e de %B, %Y")`' 
output:
  html_document:
    toc: true
    toc_depth: 2
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    toc: true
header-includes:
  - \usepackage[english]{babel}
params:
  
  file_IP: Integromics_IPmarkers.xlsx
  file_metabolome: Integromics_Metabolome.xlsx
  file_general: Integromics_1.xlsx
  file_microbiota : integromics_microbiota.xlsx
  folder.data: ../../datos
  subtitulo: Inference univariate and multivariate Metabolome
geometry: margin=2cm
---
```{r setup, include=FALSE}
require(knitr)
# include this code chunk as-is to set options
opts_chunk$set(comment = NA, prompt = TRUE, tidy = T, 
               fig.width = 7, fig.height = 7,echo = TRUE, 
               message = FALSE, warning = FALSE, cache=TRUE, fig_caption = TRUE)
Sys.setlocale("LC_TIME", "C")
```

## Load the libraries



```{r libraries}

list.of.packages <- c("xlsx","kableExtra","dplyr","ggplot2","egg","cowplot","patchwork","gridExtra","UsingR","car","lattice","ggpubr","ggbreak","GGally","reshape2","ggcorrplot","corrplot","ggrepel","factoextra","chemometrics","sparsepca","vegan","ca","ade4","pls","OmicsPLS")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])] ## check for packages not installed 
if(length(new.packages)) install.packages(new.packages) ## install packages if necessary
res<-unlist(lapply(list.of.packages, require,character.only = T)) ## load packages needed for the session
# if(any(res==F)){
  list.of.packages[which(res==F)]  ## show those package if you have troubles to install.

list.of.bioc.packages <- c("phyloseq","pcaMethods","ropls","mixOmics","limma","DESeq2","edgeR")
new.packages.bioc <- list.of.packages[!(list.of.bioc.packages %in% installed.packages()[,"Package"])]
if(length(new.packages.bioc)) BiocManager::install(new.packages.bioc)
res.bioc <- unlist(lapply(list.of.bioc.packages, require,character.only = T)) 
if(any(res.bioc==F)){
  list.of.packages[which(res.bioc==F)]  ## show those package if you have troubles to install.
}
```

# Funciones
```{r functions}


proces_bacteria <- function(datos,tipo){

  Order <- bacteria_phylum.abs$Order
  variables_in_bacteria <- general_data[general_data$Orden %in% Order , ]
  Sample<-variables_in_bacteria$Paciente


  if(tipo=="genera"){

    Order <- Order

    X <- datos[-1,-1]
    X<-apply(X,2,as.numeric)
    colnames(X) <- datos[1,2:ncol(datos)]
    rownames(X) <- Sample
    X.rel <- 100*X/rowSums(X)


  } else if(tipo=="phylum"){

    X <- datos[,-c(1:2)]
    colnames(X) <- colnames(datos)[3:ncol(datos)]
    rownames(X) <- Sample
    X.rel<-100*X/rowSums(X)
  }

  GROUP <-variables_in_bacteria$GROUP
  SEX <- variables_in_bacteria$SEX
  OBESE<- variables_in_bacteria$OBESE

  X<-as.data.frame(X)
  X.rel<-as.data.frame(X.rel)

  X$GROUP<-GROUP
  X$SEX<-SEX
  X$OBESE<-OBESE


  X.rel$GROUP<-GROUP
  X.rel$SEX<-SEX
  X.rel$OBESE<-OBESE

  return(list(abs=X,rel=X.rel))

}


low.count.removal = function(
                        data, # OTU count data frame of size n (sample) x p (OTU)
                        percent=0.01 # cutoff chosen
                        ){
    keep.otu = which(colSums(data)*100/(sum(colSums(data))) > percent)
    data.filter = data[,keep.otu]
    return(list(data.filter = data.filter, keep.otu = keep.otu))
}

low.count.removal.counts = function(
                        data, # OTU count data frame of size n (sample) x p (OTU)
                        abundance=10 # cutoff chosen
                        ){
    keep.otu <- which(colSums(data)>abundance)
    data.filter <-data[,keep.otu]
    return(list(data.filter = data.filter, keep.otu = keep.otu))
}

relative.fn <-function(data){
  
  return(100*data/rowSums(data))
  
}


common.bacteria <- function(datos_i,datos.comunes){
  

  ## retorna los indices con las variables comunes

  interaccion <- interaction(datos.comunes$GROUP,datos.comunes$OBESE)
  datos.comunes$interaccion <- interaccion
  grupo <-levels(interaccion)
  w<-vector(mode="list",length=6)
  names(w)<-levels(interaccion)
  for(i in 1:6){
   w[[names(w)[i]]]<-which(apply(as.matrix(datos_i[
    rownames(datos_i)%in%
      datos.comunes[datos.comunes$interaccion==grupo[i],"Paciente"],]),2,function(x) sum(x)==0))
  }

  w2<-vector(mode="list",length=6)
  names(w2)<-levels(interaccion)
  for(i in 1:6){
   w2[[names(w2)[i]]]<-which(apply(as.matrix(datos_i[
    rownames(datos_i)%in%
      datos.comunes[datos.comunes$interaccion==grupo[i],"Paciente"],]),2,function(x) sum(x)>0))
    }

  w3<-lapply(w2, names)
  common_variables <-Reduce(intersect,x = w3)

  return(common_variables)
}

fn.densidad <- function(datos,fctr){
  column_var=c("interaccion","variable","value")
  columnas <- colnames(datos)
  filas <- rownames(datos)
  X<-data.frame(apply(datos, 2, as.numeric))
  colnames(X)<-columnas
  rownames(X)<-filas
  X$fctr<-fctr
  X.melt <- melt(X)
  colnames(X.melt)<-column_var
  p<-ggplot(data=X.melt,aes(x=value,color=interaccion))+geom_density()
  
  return(p)
}

fn.boxplot<-function(datos,fctr){
  column_var=c("interaccion","variable","value")
  columnas <- colnames(datos)
  filas <- rownames(datos)
  X<-data.frame(apply(datos, 2, as.numeric))
  colnames(X)<-columnas
  rownames(X)<-filas
  X$fctr<-fctr
  X.melt <- melt(X)
  colnames(X.melt)<-column_var
  
  p<-ggplot(X.melt,aes(x=interaccion,y=value,fill=interaccion))+geom_boxplot()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  return(p)
}


graphPCA.methods <- function(datos,datos.common,varianzas){
  
  
  sujetos <-as.data.frame(datos)
  sexo<-datos.common$GROUP
  obesidad<-datos.common$OBESE
  colnames(sujetos)<-c("PC1","PC2")
  p1<-ggplot(sujetos,aes(x=PC1, y=PC2)) +
     geom_hline(yintercept = 0, color = "gray70") +
     geom_vline(xintercept = 0, color = "gray70") +
     geom_point(aes(color = sexo,shape=obesidad,fill=obesidad), alpha = 0.55, size = 3) +
     coord_cartesian(xlim = c(min(sujetos[,1]),max(sujetos[,1]))) +
     scale_color_manual(values=c("black","red","green"))+scale_shape_manual(values=c(17,15))+xlab(paste("PC1",round(varianzas[1],2),"%"))+ylab(paste("PC2",round(varianzas[2],2),"%"))
  return(p1)
}

```


# Cargamos los datos

## Datos clinicos

```{r}
general_data <- read.xlsx(file.path(params$folder.data,params$file_general),sheetIndex =1)
general_data$SEX <- factor(general_data$SEX, levels = c(0, 2), labels = c("Female", "Male"))
general_data$OBESE <- factor(general_data$OBESE, levels = c(0, 1), labels = c("No Obese", "Obese"))
general_data$GROUP <- factor(general_data$GROUP, levels = c(0, 1, 2), labels = c("Female",  "PCOS", "Male"))


```


## Datos IP

```{r}

IP_file<-file.path(params$folder.data,file=params$file_IP) ## file path of the data
IP.tmp<-read.xlsx(IP_file,sheetIndex = 1)
IP<-IP.tmp[,-7] # remove the missin values column
IP$SEX <- factor(IP$SEX,levels=c(0,2),labels = c("Female","Male"))
IP$OBESE <- factor(IP$OBESE,levels=c(0,1),labels = c("No Obese","Obese"))
IP$GROUP <- factor(IP$GROUP,levels=c(0,1,2),labels = c("Female","PCOS","Male"))

```

## Datos Metaboloma

```{r}
metabolome_file<-file.path(params$folder.data,file=params$file_metabolome) ## file path of the data
metabolome.tmp<-read.xlsx(metabolome_file,sheetIndex = 1) # read the data
metabolites_name <- read.xlsx(metabolome_file,sheetIndex = 2)$Metabolitos # read the names of the metabolites
any(is.na(metabolome.tmp)) # there is a row of missing values.
metabolome<-metabolome.tmp[-nrow(metabolome.tmp),] ## remove the missing value row
metabolome$GROUP<-general_data$GROUP ## add GROUP variable 
metabolome$OBESE <- general_data$OBESE ## add OBESE variables.

```


## Datos microbioma

```{r}
bacteria_phylum.abs <- as.data.frame(read.xlsx(file.path(params$folder.data,params$file_microbiota), sheetIndex = 1))
bacteria_genera.abs <- as.data.frame(t(read.xlsx(file.path(params$folder.data,params$file_microbiota), sheetIndex = 3)))
phylum.list <- proces_bacteria(bacteria_phylum.abs,tipo="phylum")
genera.list <- proces_bacteria(bacteria_genera.abs,tipo="genera")

phylum.abs <- phylum.list$abs;phylum.rel<-phylum.list$rel
genera.abs <- genera.list$abs;genera.rel<-genera.list$rel

```

# Preparando los datos para la integracion
```{r}
variables_in_bacteria <- general_data[general_data$Orden %in% bacteria_phylum.abs$Order, ]

```

## Datos Clinicos
**NOTE we are going to include also on both levels of integration body measures**
We exclude ISI because it is measured after the load of glucose, the mean values, to exclude redundancy of the data, ant the postprandial levels. 


```{r}
sujetos <- variables_in_bacteria$Paciente
Clin.tmp <- variables_in_bacteria[,-c(1,2,3,4,5)] # get rid off orden, paciente, sex, group and obese
auc_clin <- colnames(Clin.tmp)[grep("AUC",colnames(Clin.tmp))]
mean_basal_clin<-colnames(Clin.tmp)[grep("_B$",colnames(Clin.tmp))]
sog <- colnames(Clin.tmp)[grep("^SO[GLP]",colnames(Clin.tmp))]
variables_clin <- colnames(Clin.tmp)
testosterona <-colnames(Clin.tmp)[grep("TEST",colnames(Clin.tmp))]
interr<-colnames(Clin.tmp)[grep("interaccion",colnames(Clin.tmp))]
ratio <- colnames(Clin.tmp)[grep("Ratio",colnames(Clin.tmp))]
w<-c(which(variables_clin %in% auc_clin),which(variables_clin %in% mean_basal_clin),which(variables_clin %in% sog),which(variables_clin %in% testosterona),
which(variables_clin %in% interr),which(variables_clin %in% ratio))
Clin <- Clin.tmp[,-w] 
rownames(Clin) <- sujetos
colnames(Clin)
```


## Datos IP
```{r}
IP.tmp <- IP[IP$Paciente %in% variables_in_bacteria$Paciente,]
redundant_variables <- c("Orden","Paciente","SEX","GROUP","OBESE","BMI")
variables_ip <- colnames(IP.tmp)
auc_ip <- variables_ip[grep("AUC",variables_ip)]
mean_ip <- variables_ip[grep("_0$",variables_ip)]
w <- c(which(variables_ip %in% redundant_variables),which(variables_ip %in% auc_ip),which(variables_ip %in% mean_ip))
IP.basal <- IP.tmp[,-w]
rownames(IP.basal) <- sujetos

```

## Datos metaboloma

```{r}
metabolome.basal.tmp <- metabolome[metabolome$Order %in% variables_in_bacteria$Orden,]
redundant_data_meta <- c("Order","Sample","GROUP","OBESE")
variables_meta <- colnames(metabolome.basal.tmp)
mean_meta <- variables_meta[grep("_[GLP]0$",variables_meta)]
auc_meta <- variables_meta[grep("AU[CG]",variables_meta)]
w <- c(which(variables_meta %in% redundant_data_meta),which(variables_meta %in% mean_meta),which(variables_meta %in% auc_meta))
metabolome.basal <- metabolome.basal.tmp[,-w]
rownames(metabolome.basal) <- sujetos
colnames(metabolome.basal)

```


# Preprocess data according to the microbiome.

Because we have less subjects on the microbiome, and we have phylum and genera with variables with no records across all samples, we have to clean that.

Furthermore, we have to take care only the basal levels across all blocks. Also we have to clean the data on the metabolome, IP for repeated variables.

```{r}
variables_in_bacteria <- general_data[general_data$Orden %in% bacteria_phylum.abs$Order, ]
```

Because, we only take care of the GROUP, and OBESE variables on the analysis, we can get rid off this variables and store them apart in a new object. Furthermore, we have to have the same names for the rows, so let's keep it in another object.

```{r}
sujetos <- variables_in_bacteria$Paciente
GROUP <- variables_in_bacteria$GROUP
OBESE <- variables_in_bacteria$OBESE
interaccion <-with(variables_in_bacteria,interaction(GROUP,OBESE))
```

## Clinical data.

**NOTE we are going to include also on both levels of integration body measures**
We exclude ISI because it is measured after the load of glucose, the mean values, to exclude redundancy of the data, ant the postprandial levels. 

```{r}
sujetos <- variables_in_bacteria$Paciente
Clin.tmp <- variables_in_bacteria[,-c(1,2,3,4,5)] # get rid off orden, paciente, sex, group and obese
auc_clin <- colnames(Clin.tmp)[grep("AUC",colnames(Clin.tmp))]
mean_basal_clin<-colnames(Clin.tmp)[grep("_B$",colnames(Clin.tmp))]
sog <- colnames(Clin.tmp)[grep("^SO[GLP]",colnames(Clin.tmp))]
variables_clin <- colnames(Clin.tmp)
testosterona <-colnames(Clin.tmp)[grep("TEST",colnames(Clin.tmp))]
interr<-colnames(Clin.tmp)[grep("interaccion",colnames(Clin.tmp))]
ratio <- colnames(Clin.tmp)[grep("Ratio",colnames(Clin.tmp))]
w<-c(which(variables_clin %in% auc_clin),which(variables_clin %in% mean_basal_clin),which(variables_clin %in% sog),which(variables_clin %in% testosterona),
which(variables_clin %in% interr),which(variables_clin %in% ratio))
Clin <- Clin.tmp[,-w] 
rownames(Clin) <- sujetos
colnames(Clin)
```


## IP data

```{r}
IP.tmp <- IP[IP$Paciente %in% variables_in_bacteria$Paciente,]
redundant_variables <- c("Orden","Paciente","SEX","GROUP","OBESE","BMI")
variables_ip <- colnames(IP.tmp)
auc_ip <- variables_ip[grep("AUC",variables_ip)]
mean_ip <- variables_ip[grep("_0$",variables_ip)]
w <- c(which(variables_ip %in% redundant_variables),which(variables_ip %in% auc_ip),which(variables_ip %in% mean_ip))
IP.basal <- IP.tmp[,-w]
rownames(IP.basal) <- sujetos
```

## Metabolome data (37)

```{r}
metabolome.basal.tmp <- metabolome[metabolome$Order %in% variables_in_bacteria$Orden,]
redundant_data_meta <- c("Order","Sample","GROUP","OBESE")
variables_meta <- colnames(metabolome.basal.tmp)
mean_meta <- variables_meta[grep("_[GLP]0$",variables_meta)]
auc_meta <- variables_meta[grep("AU[CG]",variables_meta)]
w <- c(which(variables_meta %in% redundant_data_meta),which(variables_meta %in% mean_meta),which(variables_meta %in% auc_meta))
metabolome.basal_mean <- metabolome.basal.tmp[,-w]
rownames(metabolome.basal_mean) <- sujetos
colnames(metabolome.basal_mean)
```

## Metabolome data (todos)

```{r}
metabolome.basal.tmp <- metabolome[metabolome$Order %in% variables_in_bacteria$Orden,]
redundant_data_meta <- c("Order","Sample","GROUP","OBESE")
variables_meta <- colnames(metabolome.basal.tmp)
mean_meta <- variables_meta[grep("_[GLP]0$",variables_meta)]
auc_meta <- variables_meta[grep("AU[CG]",variables_meta)]
mean_meta2 <- variables_meta[grep("_mean$",variables_meta)]

w <- c(which(variables_meta %in% redundant_data_meta),which(variables_meta %in% auc_meta),which(variables_meta %in% mean_meta2))
metabolome.basal_all <- metabolome.basal.tmp[,-w]
rownames(metabolome.basal_all) <- sujetos
colnames(metabolome.basal_all)
```


# Datos que necesitamos

```{r}

pacientes <- variables_in_bacteria$Paciente
grupo<-variables_in_bacteria$GROUP
obesos <- factor(gsub(" ",".",variables_in_bacteria$OBESE))
interaccion<-factor(paste(obesos,grupo,sep="_"))
datos.common <- variables_in_bacteria[,c("GROUP","OBESE","Paciente")]
columnData <- data.frame(interaccion=interaccion)
rownames(columnData) <- pacientes

```

# Probamos con la media basales


```{r}
//
```

### Probamos aproximacion limma

```{r}
met37<-t(metabolome.basal_mean)
met37_norm <- met37/rowSums(met37)
met37_log <- apply(met37, 1,log10)
X<-((voom(met37,normalize.method = "quantile")$E))

fn.densidad(t(X),interaccion)
pcx <- prcomp(t(X),scale=F,center=T)
varianzas <- 100*(pcx$sdev^2/sum(pcx$sdev^2))
graphPCA.methods(pcx$x,datos.common,varianzas)

pcx.cor <- princomp(x=cor((X),method = "spearman"),cor=T)
varianzas <- 100*(pcx.cor$sdev^2/sum(pcx.cor$sdev^2))
graphPCA.methods(pcx.cor$scores[,1:2],datos.common,varianzas)

plsda.met <- plsda(t(X),interaccion,scale = T)
graphPCA.methods(plsda.met$variates$X,datos.common,varianzas)


euc <- vegdist(t(X),method = "euclidean")
graphPCA.methods(as.matrix(euc),datos.common,varianzas)

```



```{r}
library(MetaboDiff)
```


```{r}
met<-scale(metabolome.basal)
met <- scale((log2(metabolome.basal+1)))
sse<-SummarizedExperiment(assays = t(met),colData = columnData)
fn.densidad(t(assay(sse)),interaccion)
pcx <- prcomp(met,scale=F,center=T)
varianzas <- 100*(pcx$sdev^2/sum(pcx$sdev^2))
graphPCA.methods(pcx$x,datos.common,varianzas)

pcx.cor <- princomp(x=cor(t(met),method = "spearman"),cor=T)
varianzas <- 100*(pcx.cor$sdev^2/sum(pcx.cor$sdev^2))
graphPCA.methods(pcx.cor$scores[,1:2],datos.common,varianzas)

euc <- vegdist((met),method = "euclidean")
graphPCA.methods(as.matrix(euc),datos.common,varianzas)


```


```{r}
met <- create_mae(t(metabolome.basal),rowData = data.frame(metabolitos=colnames(metabolome.basal)),colData = columnData)
(met = knn_impute(met,cutoff=0.4))
(met <- normalize_met(met))
met.trans <- (t(assays(met@ExperimentList@listData$norm_imputed)[[1]]))
fn.densidad(met.trans,interaccion)
pcx <- prcomp(met.trans,scale=T,center=T)
varianzas <- 100*(pcx$sdev^2/sum(pcx$sdev^2))
graphPCA.methods(pcx$x,datos.common,varianzas)

pcx.cor <- princomp(x=cor(t(met.trans),method = "spearman"),cor=T)
varianzas <- 100*(pcx.cor$sdev^2/sum(pcx.cor$sdev^2))
graphPCA.methods(pcx.cor$scores[,1:2],datos.common,varianzas)

euc <- vegdist(scale((met.trans)),method = "euclidean")
graphPCA.methods(as.matrix(euc),datos.common,varianzas)

```

```{r}

met3 <- 

met2 <- voom(t(metabolome.basal),plot = T)

fn.densidad(t(met2$E),interaccion)
pcx <- prcomp(met.trans,scale=T,center=T)
varianzas <- 100*(pcx$sdev^2/sum(pcx$sdev^2))
graphPCA.methods(pcx$x,datos.common,varianzas)

pcx.cor <- princomp(x=cor(t(met.trans),method = "spearman"),cor=T)
varianzas <- 100*(pcx.cor$sdev^2/sum(pcx.cor$sdev^2))
graphPCA.methods(pcx.cor$scores[,1:2],datos.common,varianzas)

euc <- vegdist(scale((met.trans)),method = "euclidean")
graphPCA.methods(as.matrix(euc),datos.common,varianzas)

```

