---
title: Integration of omics and clinical data of the hormonal, metabolic, inflammatory,
  and oxidative response in the different macronutrients of the diet
author: "Edmond Geraud"
subtitle: '`r params$subtitulo`'
date: '`r format(Sys.Date(),"%e de %B, %Y")`' 
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    toc: true
  html_document:
    toc: true
    toc_depth: 2
header-includes:
  - \usepackage[english]{babel}
params:
  file_IP: Integromics_IPmarkers.xlsx
  file_metabolome: Integromics_Metabolome.xlsx
  file_general: Integromics_1.xlsx
  file_microbiota : integromics_microbiota.xlsx
  folder.data: ../../datos
  
  subtitulo: Integration of all data sets at basal level.
geometry: margin=2cm
---

```{r setup, include=FALSE}
require(knitr)
# include this code chunk as-is to set options
opts_chunk$set(comment = NA, prompt = TRUE, tidy = T, 
               fig.width = 7, fig.height = 7,echo = TRUE, 
               message = FALSE, warning = FALSE, cache=TRUE, fig_caption = TRUE)
Sys.setlocale("LC_TIME", "C")
```

## Load the libraries

```{r libraries}

list.of.packages <- c("xlsx","kableExtra","dplyr","ggplot2","egg","cowplot","patchwork","gridExtra","UsingR","car","lattice","ggpubr","ggbreak","GGally","reshape2","ggcorrplot","corrplot","rela","ggrepel","factoextra","chemometrics","sparsepca","vegan","ca","ade4","pls","OmicsPLS")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])] ## check for packages not installed 
if(length(new.packages)) install.packages(new.packages) ## install packages if necessary
res<-unlist(lapply(list.of.packages, require,character.only = T)) ## load packages needed for the session
if(any(res==F)){
  list.of.packages[which(res==F)]  ## show those package if you have troubles to install.
}

list.of.bioc.packages <- c("phyloseq","pcaMethods","ropls","mixOmics")
new.packages.bioc <- list.of.packages[!(list.of.bioc.packages %in% installed.packages()[,"Package"])]
if(length(new.packages.bioc)) BiocManager::install(new.packages.bioc)
res.bioc <- unlist(lapply(list.of.bioc.packages, require,character.only = T)) 
if(any(res.bioc==F)){
  list.of.packages[which(res.bioc==F)]  ## show those package if you have troubles to install.
}
```


# Functions for loading data (microbiome)

```{r}
proces_bacteria <- function(datos,tipo){

	Order <- bacteria_phylum.abs$Order
	variables_in_bacteria <- general_data[general_data$Orden %in% Order , ]
	Sample<-variables_in_bacteria$Paciente


	if(tipo=="genera"){

		Order <- Order

		X <- datos[-1,-1]
		X<-apply(X,2,as.numeric)
		colnames(X) <- datos[1,2:ncol(datos)]
		rownames(X) <- Sample
		X.rel <- 100*X/rowSums(X)


	} else if(tipo=="phylum"){

		X <- datos[,-c(1:2)]
		colnames(X) <- colnames(datos)[3:ncol(datos)]
		rownames(X) <- Sample
		X.rel<-100*X/rowSums(X)
	}

	GROUP <-variables_in_bacteria$GROUP
	SEX <- variables_in_bacteria$SEX
	OBESE<- variables_in_bacteria$OBESE

	X<-as.data.frame(X)
	X.rel<-as.data.frame(X.rel)

	X$GROUP<-GROUP
	X$SEX<-SEX
	X$OBESE<-OBESE


	X.rel$GROUP<-GROUP
	X.rel$SEX<-SEX
	X.rel$OBESE<-OBESE

	return(list(abs=X,rel=X.rel))

}

```


# Load the data

## Clinical data.

```{r}
general_data <- read.xlsx(file.path(params$folder.data,params$file_general),sheetIndex =1)
general_data$SEX <- factor(general_data$SEX, levels = c(0, 2), labels = c("Female", "Male"))
general_data$OBESE <- factor(general_data$OBESE, levels = c(0, 1), labels = c("No Obese", "Obese"))
general_data$GROUP <- factor(general_data$GROUP, levels = c(0, 1, 2), labels = c("Female",  "PCOS", "Male"))
```

## IP data.

```{r}
IP_file<-file.path(params$folder.data,file=params$file_IP) ## file path of the data
IP.tmp<-read.xlsx(IP_file,sheetIndex = 1)
IP<-IP.tmp[,-7] # remove the missin values column
IP$SEX <- factor(IP$SEX,levels=c(0,2),labels = c("Female","Male"))
IP$OBESE <- factor(IP$OBESE,levels=c(0,1),labels = c("No Obese","Obese"))
IP$GROUP <- factor(IP$GROUP,levels=c(0,1,2),labels = c("Female","PCOS","Male"))
```


## Metabolome data

```{r}
metabolome_file<-file.path(params$folder.data,file=params$file_metabolome) ## file path of the data
metabolome.tmp<-read.xlsx(metabolome_file,sheetIndex = 1) # read the data
metabolites_name <- read.xlsx(metabolome_file,sheetIndex = 2)$Metabolitos # read the names of the metabolites
any(is.na(metabolome.tmp)) # there is a row of missing values.
metabolome<-metabolome.tmp[-nrow(metabolome.tmp),] ## remove the missing value row
metabolome$GROUP<-general_data$GROUP ## add GROUP variable 
metabolome$OBESE <- general_data$OBESE ## add OBESE variables.
```

## Microbiome data.

```{r}
bacteria_phylum.abs <- as.data.frame(read.xlsx(file.path(params$folder.data,params$file_microbiota), sheetIndex = 1))
bacteria_genera.abs <- as.data.frame(t(read.xlsx(file.path(params$folder.data,params$file_microbiota), sheetIndex = 3)))
phylum.list <- proces_bacteria(bacteria_phylum.abs,tipo="phylum")
genera.list <- proces_bacteria(bacteria_genera.abs,tipo="genera")

phylum.abs <- phylum.list$abs;phylum.rel<-phylum.list$rel
genera.abs <- genera.list$abs;genera.rel<-genera.list$rel
```


# Preprocess data according to the microbiome.

Because we have less subjects on the microbiome, and we have phylum and genera with variables with no records across all samples, we have to clean that.

Furthermore, we have to take care only the basal levels across all blocks. Also we have to clean the data on the metabolome, IP for repeated variables.

```{r}
variables_in_bacteria <- general_data[general_data$Orden %in% bacteria_phylum.abs$Order, ]
```

Because, we only take care of the GROUP, and OBESE variables on the analysis, we can get rid off this variables and store them apart in a new object. Furthermore, we have to have the same names for the rows, so let's keep it in another object.

```{r}
sujetos <- variables_in_bacteria$Paciente
GROUP <- variables_in_bacteria$GROUP
OBESE <- variables_in_bacteria$OBESE
interaccion <-with(variables_in_bacteria,interaction(GROUP,OBESE))
```

## Clinical data.

**NOTE we are going to include also on both levels of integration body measures**
We exclude ISI because it is measured after the load of glucose, the mean values, to exclude redundancy of the data, ant the postprandial levels. 

```{r}
Clin.tmp <- variables_in_bacteria[,-c(1,2,3,4,5)] # get rid off orden, paciente, sex, group and obese
auc_clin <- colnames(Clin.tmp)[grep("AUC",colnames(Clin.tmp))]
mean_basal_clin<-colnames(Clin.tmp)[grep("mean",colnames(Clin.tmp))]
isi <- colnames(Clin.tmp)[grep("ISI",colnames(Clin.tmp))]
variables_clin <- colnames(Clin.tmp)
w<-c(which(variables_clin %in% auc_clin),which(variables_clin %in% mean_basal_clin),which(variables_clin %in% isi))
Clin <- Clin.tmp[,-w]
rownames(Clin) <- sujetos
```


## IP data

```{r}
IP.tmp <- IP[IP$Paciente %in% variables_in_bacteria$Paciente,]
redundant_variables <- c("Orden","Paciente","SEX","GROUP","OBESE","BMI")
variables_ip <- colnames(IP.tmp)
auc_ip <- variables_ip[grep("AUC",variables_ip)]
mean_ip <- variables_ip[grep("mean",variables_ip)]
w <- c(which(variables_ip %in% redundant_variables),which(variables_ip %in% auc_ip),which(variables_ip %in% mean_ip))
IP.basal <- IP.tmp[,-w]
rownames(IP.basal) <- sujetos

```

## Metabolome data.

```{r}
metabolome.basal.tmp <- metabolome[metabolome$Order %in% variables_in_bacteria$Orden,]
redundant_data_meta <- c("Order","Sample","GROUP","OBESE")
variables_meta <- colnames(metabolome.basal.tmp)
mean_meta <- variables_meta[grep("mean",variables_meta)]
auc_meta <- variables_meta[grep("AU[CG]",variables_meta)]
w <- c(which(variables_meta %in% redundant_data_meta),which(variables_meta %in% mean_meta),which(variables_meta %in% auc_meta))
metabolome.basal <- metabolome.basal.tmp[,-w]
rownames(metabolome.basal) <- sujetos
```

## Microbiome data.

### Phylum.

```{r}
phylum.abs.tmp <- phylum.abs
redundant_data_ph <- c("GROUP","SEX","OBESE")
phylum.tmp <- phylum.abs.tmp[,-c(ncol(phylum.abs.tmp)-2,ncol(phylum.abs.tmp)-1,ncol(phylum.abs.tmp))]
phylum <- phylum.tmp[,colSums(phylum.tmp)>0]
rownames(phylum) <- sujetos
```

### Genus

```{r}
genera.abs.tmp <- genera.abs
redundant_data_ph <- c("GROUP","SEX","OBESE")
genus.tmp.tmp <- genera.abs.tmp[,-c(ncol(genera.abs.tmp)-2,ncol(genera.abs.tmp)-1,ncol(genera.abs.tmp))]
genus <- genus.tmp.tmp[,colSums(genus.tmp.tmp)>0]
rownames(genus) <- sujetos
```


# INTEGRATIO METABOLOME-MICROBIOME (GENUS ABS)

## Determime the relationship between data-sets
Firstly we could perform a pls regression to undersntand properly the correlation between both datas-sets.

```{r}
datos <- list(metabolome = metabolome.basal,
              genus = genus)
res1.pls.met_micro <- mixOmics::pls(datos$metabolome,datos$genus,ncomp=1)
rho<-cor(res1.pls.met_micro$variates$X,res1.pls.met_micro$variates$Y)
```
Thus, the design should be

```{r}
design = matrix(rho, ncol = length(datos), nrow = length(datos),
                dimnames = list(names(datos), names(datos)))
diag(design) = 0
design
```

We are going to keep all variables since we do not have the power.



```{r}

diablo.integro<- block.plsda(datos,Y=interaccion, ncomp = 5, design = design)

set.seed(123) # For reproducibility, remove for your analyses
perf.diablo.integro = perf(diablo.integro, validation = 'Mfold', folds = 10, nrepeat = 10)

#perf.diablo.tcga$error.rate  # Lists the different types of error rates

# Plot of the error rates based on weighted vote
plot(perf.diablo.integro)
```


```{r}
ncomp <- perf.diablo.integro$choice.ncomp$WeightedVote["Overall.BER", "centroids.dist"]
sgccda.res_met_gen = block.splsda(X = datos, Y = interaccion, ncomp = ncomp, design = design)
```



```{r}
plotIndiv(sgccda.res_met_gen, ind.names = FALSE, legend = TRUE, title = 'DIABLO')

```

```{r}
fn.plot <- function(X,ncomp1,ncomp2,title){
  
  grupos<- list(GROUP,OBESE)
  Group <- grupos[[1]]
  Group2<-grupos[[2]]
  
  X.df<-as.data.frame(X)
  colnames(X.df) <- paste0("PC",1:dim(X.df)[2])
  colx <- colnames(X.df)
  colores<-1:length(levels(Group))
  p1 <- ggplot(X.df,aes_string(x=colx[ncomp1],y=colx[ncomp2]))+
    theme_article()+
    geom_hline(yintercept = 0, color = "gray70") +
    geom_vline(xintercept = 0, color = "gray70") +
    geom_point(aes(color = Group,shape=Group2), alpha = 0.55, size = 3)+
    scale_fill_discrete(name = "Group") + labs(x =colx[ncomp1], y = colx[ncomp2])+
    scale_color_manual(values=colores)+ggtitle(title)
  
  return(p1)
}


plot_integration <- function(lista.variates){
  
  nombres <- names(lista.variates)
  graficos.lista <- vector(mode="list",length=length(lista.variates))
  for(i in 1:length(lista.variates)){
    
    graficos.lista[[i]]<-fn.plot(lista.variates[[i]],1,2,title=nombres[i])
  }
  
  do.call("ggarrange",c(graficos.lista,common.legend = T,legend="bottom"))
  }
```



```{r}
fn.plot. <- function(X,ncomp1,ncomp2,title){
  
  grupos<- list(GROUP,OBESE)
  Group <- grupos[[1]]
  Group2<-grupos[[2]]
  
  X.df<-as.data.frame(X)
  colnames(X.df) <- paste0("PC",1:dim(X.df)[2])
  colx <- colnames(X.df)
  colores<-1:length(levels(Group))
  p1 <- ggplot(X.df,aes_string(x=colx[ncomp1],y=colx[ncomp2]))+
    theme_article()+
    geom_hline(yintercept = 0, color = "gray70") +
    geom_vline(xintercept = 0, color = "gray70") +
    geom_point(aes(color = Group,shape=Group2), alpha = 0.55, size = 3)+
    scale_fill_discrete(name = "Group") + labs(x =colx[ncomp1], y = colx[ncomp2])+
    scale_color_manual(values=colores)+ggtitle(title)
  
  return(p1)
}


plot_integration <- function(lista.variates){
  
  nombres <- names(lista.variates)
  graficos.lista <- vector(mode="list",length=length(lista.variates))
  for(i in 1:length(lista.variates)){
    
    graficos.lista[[i]]<-fn.plot(lista.variates[[i]],1,2,title=nombres[i])
  }
  
  do.call("ggarrange",c(graficos.lista,common.legend = T,legend="bottom"))
  }
```




```{r}
fn.plot_all <- function(obj,what,ncomp1,ncomp2){
  grupos<- list(GROUP,OBESE)
  Group <- grupos[[1]]
  Group2<-grupos[[2]]
  
  datos <- obj[what][[1]]
  nombres_datos <- obj$names$colnames[1:(length(obj$names$colnames)-1)]
  n <- length(nombres_datos)
  lista_graficos <- vector(mode="list",length=n)
  names(lista_graficos) <- nombres_datos
  print(str(nombres_datos[["metabolome"]]))
  if(what=="loadings"){
  for(i in 1:n){
    X <- datos[i]
    X.df<-as.data.frame(X)
    colnames(X.df) <- paste0("PC",1:dim(X.df)[2])
    colx <- colnames(X.df)
    colores<-1:length(levels(Group))
    lista_graficos[[i]]<- ggplot(X.df,aes_string(x=colx[ncomp1],y=colx[ncomp2],label=nombres_datos[[i]]))+
      theme_article()+
      geom_hline(yintercept = 0, color = "gray70") +
      geom_vline(xintercept = 0, color = "gray70") +
      geom_point(aes(color = "red"), alpha = 0.55, size = 3)+
      scale_fill_discrete(name = "Group")+ggtitle(nombres_datos[i])+
      labs(x=paste(colx[ncomp1],round(100*obj$prop_expl_var[nombres_datos][i][ncomp1],2),"%"),                                                                        y = paste(colx[ncomp2],round(100*obj$prop_expl_var[nombres_datos][i][ncomp2],2),"%"))

    
  }
  }else if(what=="scores")
  for(i in 1:n){
    
    X <- datos[i]
    X.df<-as.data.frame(X)
    colnames(X.df) <- paste0("PC",1:dim(X.df)[2])
    colx <- colnames(X.df)
    colores<-1:length(levels(Group))
    lista_graficos <- ggplot(X.df,aes_string(x=colx[ncomp1],y=colx[ncomp2]))+
      theme_article()+
      geom_hline(yintercept = 0, color = "gray70") +
      geom_vline(xintercept = 0, color = "gray70") +
      geom_point(aes(color = Group,shape=Group2), alpha = 0.55, size = 3)+
      scale_fill_discrete(name = "Group") + labs(x =colx[ncomp1], y = colx[ncomp2])+
      scale_color_manual(values=colores)+ggtitle(nombres_datos[i])+labs(x=paste(colx[ncomp1],round(100*obj$prop_expl_var[nombres_datos][[1]][ncomp1],2),"%"),
                                                                        y = paste(colx[ncomp2],round(100*obj$prop_expl_var[nombres_datos][[1]][ncomp2],2),"%"))
    
  }
  
  return(lista_graficos)
}

```


```{r}
fn.plot_all(sgccda.res_met_gen,"loadings",1,2)

```
```{r}
plotVar(sgccda.res_met_gen,overlap = F)

```


```{r}

lista.variates.1 <-list(Metabolome = sgccda.res_met_gen$variates$metabolome,
                      Microbiome = sgccda.res_met_gen$variates$genus)

plot_integration(lista.variates.1)
```


```{r}
fn.pairs <- function(X,Y,namex,namey,ncomp1,ncomp2){
  
  grupos<- list(GROUP,OBESE)
  Group <- grupos[[1]]
  Group2<-grupos[[2]]
  
  X.df<-as.data.frame(X)
  colnames(X.df) <- paste0("PC_",namex,1:dim(X.df)[2])
  colx <- colnames(X.df)
  Y.df <- as.data.frame(Y)
  colnames(Y.df) <- paste0("PC_",namey,1:dim(Y.df)[2])
  coly <- colnames(Y.df)
  datos_conjuntos <- cbind(X.df,Y.df)
  colores<-1:length(levels(Group))
  p1 <- ggplot(datos_conjuntos,aes_string(x=colx[ncomp1],y=coly[ncomp2]))+
    theme_article()+theme(legend.position = "none")+
    geom_hline(yintercept = 0, color = "gray70") +
    geom_vline(xintercept = 0, color = "gray70") +
    geom_point(aes(color = Group,shape=Group2), alpha = 0.55, size = 3)+
    scale_fill_discrete(name = "Group") + ggtitle(paste(namex,namey,sep="-"))+
    scale_color_manual(values=colores)
  
  return(p1)
}



fun_plot_comb <- function(lista.variates){
  
  
  comb <- t(combn(names(lista.variates),2))
  lista.graficos <- vector(mode="list",length=dim(comb)[1])
  lista.graficos2 <- vector(mode="list",length=dim(comb)[1])
  for(i in 1:dim(comb)[1]){
  comb_i <- comb[i,]
  X<-lista.variates[[comb_i[1]]]
  Y <- lista.variates[[comb_i[2]]]
  lista.graficos[[i]] <- fn.pairs(X,Y,comb_i[1],comb_i[2],1,2)
  lista.graficos2[[i]] <- fn.pairs(X,Y,comb_i[1],comb_i[2],2,1)

  
  
  }
  
  print(dim(comb))
  if(dim(comb)[1]<=2){
    
      p1<-do.call("ggarrange", c(list(lista.graficos[[1]],lista.graficos2[[1]]), common.legend=T,legend="bottom"))

    return(p1)
  }else{
  p1<-do.call("ggarrange", c(lista.graficos, common.legend=T,legend="bottom"))

  p2<-do.call("ggarrange", c(lista.graficos2, common.legend=T,legend="bottom"))
  return(list(p1,p2))
  }

}
```




```{r}


lista.variates.1 <-list(Metabolome = sgccda.res_met_gen$variates$metabolome,
                      Microbiome = sgccda.res_met_gen$variates$genus)
fun_plot_comb(lista.variates.1)
```


```{r}


```




# Proper integration metabolome-microbiome. (genus abs)

## Determime the relationship between data-sets
Firstly we could perform a pls regression to undersntand properly the correlation between both datas-sets.

```{r}
datos <- list(metabolome = metabolome.basal,
              genus = genus)
res1.pls.met_micro <- mixOmics::pls(datos$metabolome,datos$genus,ncomp=1)
rho<-cor(res1.pls.met_micro$variates$X,res1.pls.met_micro$variates$Y)
```
Thus, the design should be

```{r}
design = matrix(rho, ncol = length(datos), nrow = length(datos),
                dimnames = list(names(datos), names(datos)))
diag(design) = 0
design
```

## Number of components.


```{r}

diablo.integro<- block.plsda(datos,Y=interaccion, ncomp = 5, design = design)

set.seed(123) # For reproducibility, remove for your analyses
perf.diablo.integro = perf(diablo.integro, validation = 'Mfold', folds = 10, nrepeat = 10)

#perf.diablo.tcga$error.rate  # Lists the different types of error rates

# Plot of the error rates based on weighted vote
plot(perf.diablo.integro)
```


```{r}
ncomp <- perf.diablo.integro$choice.ncomp$WeightedVote["Overall.BER", "centroids.dist"]

```

## Number of variables to select.

```{r}
BPPARAM <- BiocParallel::MulticoreParam(workers = parallel::detectCores()-1)
test.keepX <- list(metabolome = c(5:9, seq(10, 25, 5)),
                   genus = c(5:9, seq(10, 25, 5)))


tune.diablo.met_gen <- tune.block.splsda(datos, interaccion, ncomp = ncomp,
                              test.keepX = test.keepX, design = design,
                              validation = 'Mfold', folds = 5, nrepeat = 2,
                              dist = "centroids.dist",BPPARAM = BPPARAM)


```

```{r}
list.keepX = tune.diablo.met_gen$choice.keepX

```

```{r}
sgccda.res_met_gen = block.splsda(X = datos, Y = interaccion, ncomp = ncomp,
                          keepX = list.keepX, design = design)
```


```{r}
plotDiablo(sgccda.res_met_gen, ncomp = 1)

```

```{r}
plotIndiv(sgccda.res_met_gen, ind.names = FALSE, legend = TRUE, title = 'DIABLO')

```


```{r}
lista.variates.1 <-list(meta = sgccda.res_met_gen$variates$metabolome,
                      genus = sgccda.res_met_gen$variates$genus)
fun_plot_comb(lista.variates.1)
```

# TOTAL INTEGRATIVE ANALYSIS AT BASAL LEVEL.



## Blocks with Genus absolute data, Scale=T


Firstly let's investigate the relationship between blocks.

```{r}

datos <- list(clinical_data = Clin,
              IP =IP.basal,
              metabolome = metabolome.basal,
              genus = genus)

res1.pls.met_micro <- mixOmics::pls(datos$metabolome,datos$genus,ncomp=1)
(rho_met_micro <- cor(res1.pls.met_micro$variates$X,res1.pls.met_micro$variates$Y))

res1.pls.met_IP <- mixOmics::pls(datos$IP,datos$metabolome,ncomp = 1)
(rho_met_ip <- cor(res1.pls.met_IP$variates$X,res1.pls.met_IP$variates$Y))

res1.pls.met_clin <- mixOmics::pls(datos$metabolome,datos$clinical_data,ncomp=1)
(rho_pls_met_clin <- cor(res1.pls.met_clin$variates$X,res1.pls.met_clin$variates$Y))

res1.pls_micro_clin <- mixOmics::pls(datos$genus,datos$clinical_data,ncomp=1)
(rho_pls_micro_clin <- cor(res1.pls_micro_clin$variates$X,res1.pls_micro_clin$variates$Y))

res1.pls_micro_IP <- mixOmics::pls(datos$genus,datos$IP,ncomp = 1)
(rho_pls_micro_IP <- cor(res1.pls_micro_IP$variates$X,res1.pls_micro_IP$variates$Y))

res1.pls_IP_clin <- mixOmics::pls(datos$IP,datos$clinical_data,ncomp = 1)
(rho_pls_IP_clin <- cor(res1.pls_IP_clin$variates$X,res1.pls_IP_clin$variates$Y))

design =matrix(c(0,rho_pls_IP_clin,rho_pls_met_clin,rho_pls_micro_clin,
         rho_pls_IP_clin,0,rho_met_ip,rho_pls_micro_IP,
         rho_pls_met_clin,rho_met_ip,0,rho_met_micro,
         rho_pls_micro_clin,rho_pls_micro_IP,rho_met_micro,0),ncol = 4,dimnames = list(names(datos),names(datos)))
design
```

Now let's observe the optimal number of components.

```{r}

sgccda.res_total_1 = block.splsda(X = datos, Y=interaccion, ncomp = 10,
                           design = design)

set.seed(123) # for reproducibility, only when the `cpus' argument is not used
# this code takes a couple of min to run
perf.diablo_toltal_1 = mixOmics::perf(sgccda.res_total_1, validation = 'Mfold', folds = 10, nrepeat = 10)

#perf.diablo  # lists the different outputs
plot(perf.diablo_toltal_1)
```


```{r}
perf.diablo_toltal_1$choice.ncomp$WeightedVote
(ncomp = perf.diablo_toltal_1$choice.ncomp$WeightedVote["Overall.BER", "centroids.dist"])

```



```{r}
BPPARAM <- BiocParallel::MulticoreParam(workers = parallel::detectCores()-1)
test.keepX <- list(clinical_data =c(5:9, seq(10, 25, 5)),
                   IP = c(5:9, seq(10, 25, 5)),
                   metabolome = c(5:9, seq(10, 25, 5)),
                   genus = c(5:9, seq(10, 25, 5)))


tune.diablo.2 <- tune.block.splsda(datos, interaccion, ncomp = ncomp,
                              test.keepX = test.keepX, design = design,
                              validation = 'Mfold', folds = 5, nrepeat = 2,
                              dist = "centroids.dist",BPPARAM = BPPARAM)


```

```{r}
list.keepX = tune.diablo.2$choice.keepX
# $clinical_data
# [1] 25 15  5 25 25
# 
# $IP
# [1] 10 10 10  6  6
# 
# $metabolome
# [1]  5 10 25 20  6
# 
# $genus
# [1] 25  8  5  5 20
# 
# ```
```

```{r}
sgccda.res_genus = block.splsda(X = datos, Y = interaccion, ncomp = ncomp,
                          keepX = list.keepX, design = design)
```



```{r}
plotDiablo(sgccda.res_genus, ncomp = 1)

```


```{r}
plotIndiv(sgccda.res_genus, ind.names = FALSE, legend = TRUE, title = 'DIABLO',study="global")

```

```{r}

lista.variates <-list(clin=sgccda.res_genus$variates$clinical_data,
                      ip = sgccda.res_genus$variates$IP,
                      meta = sgccda.res_genus$variates$metabolome,
                      micro = sgccda.res_genus$variates$genus)
fun_plot_comb(lista.variates)




```


# TOTAL INTEGRATION ALL BLOCKS absolute abundance

## SCALE=T abslute

```{r}
datos <- list(clinical_data = Clin,
              IP =IP.basal,
              metabolome = metabolome.basal,
              genus = genus,
              phylum = phylum)


res1.pls.met_micro <- mixOmics::pls(datos$metabolome,datos$genus,ncomp=1,scale=T)
(rho_met_micro <- cor(res1.pls.met_micro$variates$X,res1.pls.met_micro$variates$Y))

res1.pls.met_IP <- mixOmics::pls(datos$IP,datos$metabolome,ncomp = 1,scale=T)
(rho_met_ip <- cor(res1.pls.met_IP$variates$X,res1.pls.met_IP$variates$Y))

res1.pls.met_clin <- mixOmics::pls(datos$metabolome,datos$clinical_data,ncomp=1,scale=T)
(rho_pls_met_clin <- cor(res1.pls.met_clin$variates$X,res1.pls.met_clin$variates$Y))

res1.pls_micro_clin <- mixOmics::pls(datos$genus,datos$clinical_data,ncomp=1,scale=T)
(rho_pls_micro_clin <- cor(res1.pls_micro_clin$variates$X,res1.pls_micro_clin$variates$Y))

res1.pls_micro_IP <- mixOmics::pls(datos$genus,datos$IP,ncomp = 1,scale=T)
(rho_pls_micro_IP <- cor(res1.pls_micro_IP$variates$X,res1.pls_micro_IP$variates$Y))

res1.pls_IP_clin <- mixOmics::pls(datos$IP,datos$clinical_data,ncomp = 1,scale=T)
(rho_pls_IP_clin <- cor(res1.pls_IP_clin$variates$X,res1.pls_IP_clin$variates$Y))

res1.pls.clin_filo<-mixOmics::pls(datos$phylum,datos$clinical_data,ncomp = 1,scale=T)
(rho_pls_clin_filo <- cor(res1.pls.clin_filo$variates$X,res1.pls.clin_filo$variates$Y))

res1.pls.IP_filo<-mixOmics::pls(datos$IP,datos$phylum,ncomp = 1,scale=T)
(rho_pls_IP_filo <- cor(res1.pls.IP_filo$variates$X,res1.pls.IP_filo$variates$Y))


res1.pls.met_filo<-mixOmics::pls(datos$metabolome,datos$phylum,ncomp = 1,scale=T)
(rho_pls_met_filo <- cor(res1.pls.met_filo$variates$X,res1.pls.met_filo$variates$Y))

res1.pls.micro_filo<-mixOmics::pls(datos$genus,datos$phylum,ncomp = 1,scale=T)
(rho_pls_micro_filo <- cor(res1.pls.micro_filo$variates$X,res1.pls.micro_filo$variates$Y))


design =matrix(c(0,rho_pls_IP_clin,rho_pls_met_clin,rho_pls_micro_clin,rho_pls_clin_filo,
         rho_pls_IP_clin,0,rho_met_ip,rho_pls_micro_IP,rho_pls_IP_filo,
         rho_pls_met_clin,rho_met_ip,0,rho_met_micro,rho_pls_met_filo,
         rho_pls_micro_clin,rho_pls_micro_IP,rho_met_micro,0,rho_pls_micro_filo,
         rho_pls_clin_filo,rho_pls_IP_filo,rho_pls_met_filo,rho_pls_micro_filo,0),ncol=5,dimnames = list(names(datos),names(datos)))


design

```


```{r}
sgccda.res_total = block.splsda(X = datos, Y=interaccion, ncomp = 10,
                           design = design,scale=T)

set.seed(123) # for reproducibility, only when the `cpus' argument is not used
# this code takes a couple of min to run
perf.diablo_total = mixOmics::perf(sgccda.res_total, validation = 'Mfold', folds = 10, nrepeat = 10,scale=F)

#perf.diablo  # lists the different outputs
plot(perf.diablo_total)
```




```{r}
perf.diablo_total$choice.ncomp$WeightedVote
(ncomp = perf.diablo_total$choice.ncomp$WeightedVote["Overall.BER", "centroids.dist"])

```



```{r}
BPPARAM <- BiocParallel::MulticoreParam(workers = parallel::detectCores()-1)
test.keepX <- list(clinical_data =c(5:9, seq(10, 25, 5)),
                   IP = c(5:9, seq(10, 25, 5)),
                   metabolome = c(5:9, seq(10, 25, 5)),
                   phylum = c(5:9, seq(10, 25, 5)),
                   genus = c(5:9, seq(10, 25, 5)))


tune.diablo.3 <- tune.block.splsda(datos, interaccion, ncomp = ncomp,
                              test.keepX = test.keepX, design = design,
                              validation = 'Mfold', folds = 5, nrepeat = 2,
                              dist = "centroids.dist",BPPARAM = BPPARAM)


```

```{r}
list.keepX = tune.diablo.3$choice.keepX

```

```{r}
sgccda.res_total = block.splsda(X = datos, Y = interaccion, ncomp = ncomp,
                          keepX = list.keepX, design = design)
```
```{r}
plotDiablo(sgccda.res_total, ncomp = 1)

```



```{r}
  plotIndiv(sgccda.res_total, ind.names = FALSE, legend = TRUE, title = 'DIABLO',study="global")

```

```{r}
lista.variates.5 <-list(clin=sgccda.res_total$variates$clinical_data,
                      ip = sgccda.res_total$variates$IP,
                      meta = sgccda.res_total$variates$metabolome,
                      genus = sgccda.res_total$variates$genus,
                      filo = sgccda.res_total$variates$phylum)


fun_plot_comb(lista.variates.5)

```
