---
title: Integration of omics and clinical data of the hormonal, metabolic, inflammatory,
  and oxidative response in the different macronutrients of the diet
author: "Edmond Geraud"
subtitle: '`r params$subtitulo`'
date: '`r format(Sys.Date(),"%e de %B, %Y")`' 
output:
  html_document:
    toc: true
    toc_depth: 2
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    toc: true
header-includes:
  - \usepackage[english]{babel}
params:
  
  file_IP: Integromics_IPmarkers.xlsx
  file_metabolome: Integromics_Metabolome.xlsx
  file_general: Integromics_1.xlsx
  file_microbiota : integromics_microbiota.xlsx
  folder.data: ../../datos
  subtitulo: Inference univariate and multivariate microbiome
geometry: margin=2cm
---
```{r setup, include=FALSE}
require(knitr)
# include this code chunk as-is to set options
opts_chunk$set(comment = NA, prompt = TRUE, tidy = T, 
               fig.width = 7, fig.height = 7,echo = TRUE, 
               message = FALSE, warning = FALSE, cache=TRUE, fig_caption = TRUE)
Sys.setlocale("LC_TIME", "C")
```

## Load the libraries



```{r libraries}

list.of.packages <- c("xlsx","kableExtra","dplyr","ggplot2","egg","cowplot","patchwork","gridExtra","UsingR","car","lattice","ggpubr","ggbreak","GGally","reshape2","ggcorrplot","corrplot","ggrepel","factoextra","chemometrics","sparsepca","vegan","ca","ade4","pls","OmicsPLS","Hotelling")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])] ## check for packages not installed 
if(length(new.packages)) install.packages(new.packages) ## install packages if necessary
res<-unlist(lapply(list.of.packages, require,character.only = T)) ## load packages needed for the session
# if(any(res==F)){
  list.of.packages[which(res==F)]  ## show those package if you have troubles to install.

list.of.bioc.packages <- c("phyloseq","pcaMethods","ropls","mixOmics","limma","DESeq2","edgeR")
new.packages.bioc <- list.of.packages[!(list.of.bioc.packages %in% installed.packages()[,"Package"])]
if(length(new.packages.bioc)) BiocManager::install(new.packages.bioc)
res.bioc <- unlist(lapply(list.of.bioc.packages, require,character.only = T)) 
if(any(res.bioc==F)){
  list.of.packages[which(res.bioc==F)]  ## show those package if you have troubles to install.
}
```

# Funciones
```{r functions}


proces_bacteria <- function(datos,tipo){

  Order <- bacteria_phylum.abs$Order
  variables_in_bacteria <- general_data[general_data$Orden %in% Order , ]
  Sample<-variables_in_bacteria$Paciente


  if(tipo=="genera"){

    Order <- Order

    X <- datos[-1,-1]
    X<-apply(X,2,as.numeric)
    colnames(X) <- datos[1,2:ncol(datos)]
    rownames(X) <- Sample
    X.rel <- 100*X/rowSums(X)


  } else if(tipo=="phylum"){

    X <- datos[,-c(1:2)]
    colnames(X) <- colnames(datos)[3:ncol(datos)]
    rownames(X) <- Sample
    X.rel<-100*X/rowSums(X)
  }

  GROUP <-variables_in_bacteria$GROUP
  SEX <- variables_in_bacteria$SEX
  OBESE<- variables_in_bacteria$OBESE

  X<-as.data.frame(X)
  X.rel<-as.data.frame(X.rel)

  X$GROUP<-GROUP
  X$SEX<-SEX
  X$OBESE<-OBESE


  X.rel$GROUP<-GROUP
  X.rel$SEX<-SEX
  X.rel$OBESE<-OBESE

  return(list(abs=X,rel=X.rel))

}


low.count.removal = function(
                        data, # OTU count data frame of size n (sample) x p (OTU)
                        percent=0.01 # cutoff chosen
                        ){
    keep.otu = which(colSums(data)*100/(sum(colSums(data))) > percent)
    data.filter = data[,keep.otu]
    return(list(data.filter = data.filter, keep.otu = keep.otu))
}

low.count.removal.counts = function(
                        data, # OTU count data frame of size n (sample) x p (OTU)
                        abundance=10 # cutoff chosen
                        ){
    keep.otu <- which(colSums(data)>abundance)
    data.filter <-data[,keep.otu]
    return(list(data.filter = data.filter, keep.otu = keep.otu))
}

relative.fn <-function(data){
  
  return(100*data/rowSums(data))
  
}


common.bacteria <- function(datos_i,datos.comunes){
  

  ## retorna los indices con las variables comunes

  interaccion <- interaction(datos.comunes$GROUP,datos.comunes$OBESE)
  datos.comunes$interaccion <- interaccion
  grupo <-levels(interaccion)
  w<-vector(mode="list",length=6)
  names(w)<-levels(interaccion)
  for(i in 1:6){
   w[[names(w)[i]]]<-which(apply(as.matrix(datos_i[
    rownames(datos_i)%in%
      datos.comunes[datos.comunes$interaccion==grupo[i],"Paciente"],]),2,function(x) sum(x)==0))
  }

  w2<-vector(mode="list",length=6)
  names(w2)<-levels(interaccion)
  for(i in 1:6){
   w2[[names(w2)[i]]]<-which(apply(as.matrix(datos_i[
    rownames(datos_i)%in%
      datos.comunes[datos.comunes$interaccion==grupo[i],"Paciente"],]),2,function(x) sum(x)>0))
    }

  w3<-lapply(w2, names)
  common_variables <-Reduce(intersect,x = w3)

  return(common_variables)
}

fn.densidad <- function(datos,fctr){
  column_var=c("interaccion","variable","value")
  columnas <- colnames(datos)
  filas <- rownames(datos)
  X<-data.frame(apply(datos, 2, as.numeric))
  colnames(X)<-columnas
  rownames(X)<-filas
  X$fctr<-fctr
  X.melt <- melt(X)
  colnames(X.melt)<-column_var
  p<-ggplot(data=X.melt,aes(x=value,color=interaccion))+geom_density()
  
  return(p)
}

fn.boxplot<-function(datos,fctr){
  column_var=c("interaccion","variable","value")
  columnas <- colnames(datos)
  filas <- rownames(datos)
  X<-data.frame(apply(datos, 2, as.numeric))
  colnames(X)<-columnas
  rownames(X)<-filas
  X$fctr<-fctr
  X.melt <- melt(X)
  colnames(X.melt)<-column_var
  
  p<-ggplot(X.melt,aes(x=interaccion,y=value,fill=interaccion))+geom_boxplot()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  return(p)
}


graphPCA.methods <- function(datos,datos.common,varianzas){
  
  
  sujetos <-as.data.frame(datos)
  sexo<-datos.common$GROUP
  obesidad<-datos.common$OBESE
  colnames(sujetos)<-c("PC1","PC2")
  p1<-ggplot(sujetos,aes(x=PC1, y=PC2)) +
     geom_hline(yintercept = 0, color = "gray70") +
     geom_vline(xintercept = 0, color = "gray70") +
     geom_point(aes(color = sexo,shape=obesidad,fill=obesidad), alpha = 0.55, size = 3) +
     coord_cartesian(xlim = c(min(sujetos[,1]),max(sujetos[,1]))) +
     scale_color_manual(values=c("black","red","green"))+scale_shape_manual(values=c(17,15))+xlab(paste("PC1",round(varianzas[1],2),"%"))+ylab(paste("PC2",round(varianzas[2],2),"%"))
  return(p1)
}

```


# Cargamos los datos

## Datos clinicos

```{r}
general_data <- read.xlsx(file.path(params$folder.data,params$file_general),sheetIndex =1)
general_data$SEX <- factor(general_data$SEX, levels = c(0, 2), labels = c("Female", "Male"))
general_data$OBESE <- factor(general_data$OBESE, levels = c(0, 1), labels = c("No Obese", "Obese"))
general_data$GROUP <- factor(general_data$GROUP, levels = c(0, 1, 2), labels = c("Female",  "PCOS", "Male"))


```


## Datos IP

```{r}

IP_file<-file.path(params$folder.data,file=params$file_IP) ## file path of the data
IP.tmp<-read.xlsx(IP_file,sheetIndex = 1)
IP<-IP.tmp[,-7] # remove the missin values column
IP$SEX <- factor(IP$SEX,levels=c(0,2),labels = c("Female","Male"))
IP$OBESE <- factor(IP$OBESE,levels=c(0,1),labels = c("No Obese","Obese"))
IP$GROUP <- factor(IP$GROUP,levels=c(0,1,2),labels = c("Female","PCOS","Male"))

```

## Datos Metaboloma

```{r}
metabolome_file<-file.path(params$folder.data,file=params$file_metabolome) ## file path of the data
metabolome.tmp<-read.xlsx(metabolome_file,sheetIndex = 1) # read the data
metabolites_name <- read.xlsx(metabolome_file,sheetIndex = 2)$Metabolitos # read the names of the metabolites
any(is.na(metabolome.tmp)) # there is a row of missing values.
metabolome<-metabolome.tmp[-nrow(metabolome.tmp),] ## remove the missing value row
metabolome$GROUP<-general_data$GROUP ## add GROUP variable 
metabolome$OBESE <- general_data$OBESE ## add OBESE variables.

```


## Datos microbioma

```{r}
bacteria_phylum.abs <- as.data.frame(read.xlsx(file.path(params$folder.data,params$file_microbiota), sheetIndex = 1))
bacteria_genera.abs <- as.data.frame(t(read.xlsx(file.path(params$folder.data,params$file_microbiota), sheetIndex = 3)))
phylum.list <- proces_bacteria(bacteria_phylum.abs,tipo="phylum")
genera.list <- proces_bacteria(bacteria_genera.abs,tipo="genera")

phylum.abs <- phylum.list$abs;phylum.rel<-phylum.list$rel
genera.abs <- genera.list$abs;genera.rel<-genera.list$rel

```

# Preparando los datos para la integracion
```{r}
variables_in_bacteria <- general_data[general_data$Orden %in% bacteria_phylum.abs$Order, ]

```

## Datos Clinicos
**NOTE we are going to include also on both levels of integration body measures**
We exclude ISI because it is measured after the load of glucose, the mean values, to exclude redundancy of the data, ant the postprandial levels. 


```{r}
sujetos <- variables_in_bacteria$Paciente
Clin.tmp <- variables_in_bacteria[,-c(1,2,3,4,5)] # get rid off orden, paciente, sex, group and obese
auc_clin <- colnames(Clin.tmp)[grep("AUC",colnames(Clin.tmp))]
mean_basal_clin<-colnames(Clin.tmp)[grep("_B$",colnames(Clin.tmp))]
sog <- colnames(Clin.tmp)[grep("^SO[GLP]",colnames(Clin.tmp))]
variables_clin <- colnames(Clin.tmp)
testosterona <-colnames(Clin.tmp)[grep("TEST",colnames(Clin.tmp))]
interr<-colnames(Clin.tmp)[grep("interaccion",colnames(Clin.tmp))]
ratio <- colnames(Clin.tmp)[grep("Ratio",colnames(Clin.tmp))]
w<-c(which(variables_clin %in% auc_clin),which(variables_clin %in% mean_basal_clin),which(variables_clin %in% sog),which(variables_clin %in% testosterona),
which(variables_clin %in% interr),which(variables_clin %in% ratio))
Clin <- Clin.tmp[,-w] 
rownames(Clin) <- sujetos
colnames(Clin)
```


## Datos IP
```{r}
IP.tmp <- IP[IP$Paciente %in% variables_in_bacteria$Paciente,]
redundant_variables <- c("Orden","Paciente","SEX","GROUP","OBESE","BMI")
variables_ip <- colnames(IP.tmp)
auc_ip <- variables_ip[grep("AUC",variables_ip)]
mean_ip <- variables_ip[grep("_0$",variables_ip)]
w <- c(which(variables_ip %in% redundant_variables),which(variables_ip %in% auc_ip),which(variables_ip %in% mean_ip))
IP.basal <- IP.tmp[,-w]
rownames(IP.basal) <- sujetos

```

## Datos metaboloma

```{r}
metabolome.basal.tmp <- metabolome[metabolome$Order %in% variables_in_bacteria$Orden,]
redundant_data_meta <- c("Order","Sample","GROUP","OBESE")
variables_meta <- colnames(metabolome.basal.tmp)
mean_meta <- variables_meta[grep("_[GLP]0$",variables_meta)]
auc_meta <- variables_meta[grep("AU[CG]",variables_meta)]
w <- c(which(variables_meta %in% redundant_data_meta),which(variables_meta %in% mean_meta),which(variables_meta %in% auc_meta))
metabolome.basal <- metabolome.basal.tmp[,-w]
rownames(metabolome.basal) <- sujetos
colnames(metabolome.basal)

```
# Datos que necesitamos

```{r}
pacientes <- variables_in_bacteria$Paciente
```

## Datos microbioma

### Datos genero

```{r}

genera.abs.tmp <- genera.abs
redundant_data_ph <- c("GROUP","SEX","OBESE")
genus.tmp.tmp <- genera.abs.tmp[,-c(ncol(genera.abs.tmp)-2,ncol(genera.abs.tmp)-1,ncol(genera.abs.tmp))]
genus <- genus.tmp.tmp[,colSums(genus.tmp.tmp)>0]
rownames(genus) <- sujetos
## removal of genus unidentified

genus<-genus[,-grep("^ud-",colnames(genus))]

common_genus <-common.bacteria(genus,variables_in_bacteria)

genus<-genus[,common_genus]
genus.abs <- low.count.removal.counts(genus,10)$data.filter

```

# Preparando inferencia


```{r}
variables_in_bacteria$SEX <- as.factor(variables_in_bacteria$SEX)
grupo<-variables_in_bacteria$GROUP
obesidad <- factor(gsub(" ",".",variables_in_bacteria$OBESE))
interaccion<-factor(paste(obesidad,grupo,sep="_"))
sexo <- variables_in_bacteria$SEX
datos.common <- variables_in_bacteria[,c("GROUP","OBESE","Paciente")]
columnData.sexo <- data.frame(sexo=sexo)
rownames(columnData.sexo) <- pacientes
columnData.obesidad <- data.frame(obesidad=obesidad)
rownames(columnData.obesidad) <- pacientes
columnData.grupo <- data.frame(grupo=grupo)
rownames(columnData.grupo) <- pacientes
columnData.interaccion <- data.frame(interaccion=interaccion)
rownames(columnData.interaccion) <- pacientes
```

## Descriptiva de variables

# interaccion 2
#### Descriptiva de la inferenica
```{r}
columnData.interaccion <- data.frame(obesidad=obesidad,grupo=grupo,interaccion=interaccion)
rownames(columnData.interaccion)<- pacientes
dds.interaccion <- DESeqDataSetFromMatrix(t(genus.abs),colData = columnData.interaccion,design = ~interaccion)

gm_mean = function(x, na.rm=TRUE){
  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
}

geoMeans = apply(counts(dds.interaccion), 1, gm_mean)

diagdds = estimateSizeFactors(dds.interaccion, geoMeans = geoMeans)

dds.interaccion.nb.2 = DESeq(diagdds, fitType="mean")

# diagdds<-varianceStabilizingTransformation(diagdds, blind = TRUE, fitType = "local")

csa <-getVarianceStabilizedData(dds.interaccion.nb.2)
# dds.interaccion.nb.2 <- DESeq(dds.interaccion,test="Wald",sfType = "poscounts",fitType = "parametric")
# dds.interaccion.lrt.2 <- DESeq(dds.interaccion,test="LRT",sfType = "poscounts",reduced = ~1)


# datos <- t(rlog(counts(dds.interaccion.nb.2)))
datos = as.data.frame(t(csa))
p1<-fn.densidad(datos = datos,fctr = interaccion)
p2 <- fn.boxplot(datos = datos,fctr = interaccion)
pca.log <- prcomp(t(datos),center = T,scale. = T)
pca.var <- 100*(pca.log$sdev^2/sum(pca.log$sdev^2))
p3<-graphPCA.methods(pca.log$rotation[,1:2],
                     datos.common = datos.common,
                     varianzas=pca.var)+ggtitle("PCA")
pca.spearman <- princomp(cor(t(datos),method = "spearman"),fix_sign = T,cor=T)

pca.spearman.scores <- pca.spearman$scores
pca.spearman.var <- 100*(pca.spearman$sdev^2/sum(pca.spearman$sdev^2))
p4<-graphPCA.methods(pca.spearman.scores[,1:2],
                 datos.common = datos.common,
                 varianzas = pca.spearman.var)+ggtitle("PCA spearman")



mds <- cmdscale(vegdist(datos,method = "euclidean"),eig = T,add = F,k=10)
varianzas <- 100*mds$eig/sum(mds$eig)
p5<-graphPCA.methods(datos=mds$points,
                     datos.common = datos.common,
                     varianzas = varianzas)+ggtitle("PCoA")


p5<-graphPCA.methods(datos=mds$points,
                     datos.common = datos.common,
                     varianzas = varianzas)+ggtitle("PCoA")



p11<-ggarrange(p1,p2,p3,p4,common.legend = T)
annotate_figure(p11,bottom = "media ajusta y rlog")
p5


```

#### Inferencia Multivariante
Antes de todo preparamos los datos
```{r}
datos$grupo <- grupo
datos$obesidad <- obesidad
datos$interaccion <- interaccion
#### gruposde interes
females <- datos[datos$grupo=="Female",-c(95,96,97)]
males <- datos[datos$grupo=="Male",-c(95,96,97)]
pcos <- datos[datos$grupo=="PCOS",-c(95,96,97)]
obesos <- datos[datos$obesidad=="Obese",-c(95,96,97)]
No.Obesos <- datos[datos$obesidad=="No.Obese",-c(95,96,97)]
#### interacciones
females.no.obese <- datos[datos$interaccion=="No.Obese_Female",-c(95,96,97)]
females.obese <- datos[datos$interaccion=="Obese_Female",-c(95,96,97)]
males.no.obese <-  datos[datos$interaccion=="No.Obese_Male",-c(95,96,97)]
males.obese <-  datos[datos$interaccion=="Obese_Male",-c(95,96,97)]
pcos.no.obese <- datos[datos$interaccion=="No.Obese_PCOS",-c(95,96,97)]
pcos.obese <- datos[datos$interaccion=="Obese_PCOS",-c(95,96,97)]


(res<-adonis2(datos[,-c(95,96,97)]~datos$grupo+datos$obesidad+datos$grupo:datos$obesidad))
```

### Analisis post-hoc
```{r}
library(Hotelling)
```

#### Obesidad
```{r}
(res<-hotelling.test(obesos,No.Obesos,shrinkage = T,var.equal = T,perm = T,B=100))

```
### grupo
```{r}
(res<-hotelling.test(females,males,shrinkage = T,var.equal = T,perm = T,B=100))


(res<-hotelling.test(females,pcos,shrinkage = T,var.equal = T,perm = T,B=100))


(res<-hotelling.test(males,pcos,shrinkage = T,var.equal = T,perm = T,B=100))

```




#### interacciones

```{r}
(res<-hotelling.test(females.no.obese,females.obese,shrinkage = T,var.equal = T,perm = T,B=100))
(res<-hotelling.test(males.no.obese,males.obese,shrinkage = T,var.equal = T,perm = T,B=100))
(res<-hotelling.test(pcos.no.obese,pcos.obese,shrinkage = T,var.equal = T,perm = T,B=100))

```
## Univariante

```{r}
res <-results(dds.interaccion.nb.2)
res[which(res$padj<0.05),]
```


El efecto de la obesidad para las mujeres control
```{r}
res <- results(dds.interaccion.nb.2,contrast=c("obesidad","No.Obese","Obese"))
res[which(res$padj<0.05),]
```

EL efecto de la obesidad de las mujeres con PCOS comparadas con las mujeres control
```{r}
res <-results(dds.interaccion.nb.2, contrast=list( c("obesidad_Obese_vs_No.Obese","grupoPCOS.obesidadObese") ))
res[which(res$padj<0.05),]
```
El efecto de la obesidad en los hombres comparado con las mujeres control
```{r}
res <-results(dds.interaccion.nb.2, contrast=list( c("obesidad_Obese_vs_No.Obese","grupoMale.obesidadObese") ))
res[which(res$padj<0.05),]
```
La obesidad comparada es diferente? entre hombres y mujeres con el sindrome
```{r}
res <-results(dds.interaccion.nb.2, contrast=list( c("grupoPCOS.obesidadObese","grupoMale.obesidadObese") ))
res[which(res$padj<0.05),]
```
Cuales generos con diferentes entre la obesidad y entre los 3 subgrupos
```{r}
res <- results(dds.interaccion.lrt.2)
res[which(res$padj<0.05),]
```


### MI interaccion



```{r}

(combinaciones.interaccion <- matrix(c(levels(interaccion)[1:3],levels(interaccion)[4:6]),ncol=2,nrow = 3))
combinaciones.grupo <- t(combn(levels(grupo),2))
combinaciones.obesos <- t(combn(levels(obesidad),2))
lista.obesos <- list(Obese=obesos,No.Obese=No.Obesos)
lista.grupo <- list(Female=females,Male=males,PCOS=pcos)
lista.sujetos <- list(No.Obese_Female=females.no.obese,
                      Obese_Female=females.obese,
                      No.Obese_PCOS=pcos.no.obese,
                      Obese_PCOS=pcos.obese,
                      No.Obese_Male=males.no.obese,
                      Obese_Male=males.obese)

matriz.obesos <- as.data.frame(matrix(NA,ncol=1,nrow=94))
colnames(matriz.obesos)<-apply(combinaciones.obesos, 1, function(x) paste(x[1],x[2],sep="_vs_"))
rownames(matriz.obesos) <-colnames(genus.abs)

## obesos

for(j in 1:94){
for(i in 1:dim(combinaciones.obesos)[1]){
  x<-lista.obesos[[combinaciones.obesos[i,1]]][,j]
  y <- lista.obesos[[combinaciones.obesos[i,2]]][,j]
  matriz.obesos[j,i]<-wilcox.test(x,y)$p.value

  
}
}

matriz.grupo <- as.data.frame(matrix(NA,ncol=3,nrow=94))
colnames(matriz.grupo)<-apply(combinaciones.grupo, 1, function(x) paste(x[1],x[2],sep="_vs_"))
rownames(matriz.grupo) <-colnames(genus.abs)

## grupo
for(j in 1:94){
for(i in 1:dim(combinaciones.grupo)[1]){
  x<-lista.grupo[[combinaciones.grupo[i,1]]][,j]
  y <- lista.grupo[[combinaciones.grupo[i,2]]][,j]
  matriz.grupo[j,i]<-wilcox.test(x,y)$p.value

  
}
}

matriz.grupo.adj <- as.data.frame(t(apply(matriz.grupo, 1, function(x) p.adjust(x,method="BH"))))


## interaccion

matriz.interaccion <- as.data.frame(matrix(NA,ncol=3,nrow=94))
colnames(matriz.interaccion)<-apply(combinaciones.interaccion, 1, function(x) paste(x[1],x[2],sep="_vs_"))
rownames(matriz.interaccion) <-colnames(genus.abs)

## grupo
for(j in 1:94){
for(i in 1:dim(combinaciones.interaccion)[1]){
  x<-lista.sujetos[[combinaciones.interaccion[i,1]]][,j]
  y <- lista.sujetos[[combinaciones.interaccion[i,2]]][,j]
  matriz.interaccion[j,i]<-wilcox.test(x,y)$p.value

  
}
}

matriz.interaccion.adj <- as.data.frame(t(apply(matriz.interaccion, 1, function(x) p.adjust(x,method="BH"))))


# 
# 
# 
# for(j in 1:94){
# for(i in 1:dim(combinaciones)[1]){
#   print(colnames(lista.sujetos[[combinaciones[i,1]]])[j])
#   x<-lista.sujetos[[combinaciones[i,1]]][,j]
#   y <- lista.sujetos[[combinaciones[i,2]]][,j]
#   lista.pvalores[j,i]<-wilcox.test(x,y)$p.value
# 
#   
# }
# }
# 
# p.adj_valores <-as.data.frame(apply(lista.pvalores, 1,function(x) p.adjust(x,method = "BH")))
# colnames(p.adj_valores) <- rownames(genus.t)
```
## univariante a mi manera
### obesos
```{r}
df.obesos <-data.frame(obesos_vs_no.obesos=matriz.obesos[which(matriz.obesos[,1]<0.05),])
rownames(df.obesos) <- rownames(matriz.obesos)[which(matriz.obesos[,1]<0.05)]
df.obesos
```
### grupo

```{r}
w <- which(apply(matriz.grupo.adj, 1, function(x) any(x<0.05)))
df.grupo <-data.frame(matriz.grupo.adj[w,])
colnames(df.grupo) <- colnames(matriz.grupo.adj)
rownames(df.grupo) <- rownames(matriz.grupo.adj)[w]
df.grupo

```

```{r}
intersect(rownames(df.grupo),rownames(df.obesos))
```

## interaccion mia

```{r}
w <- which(apply(matriz.interaccion.adj, 1, function(x) any(x<0.05)))
df.interaccion <-data.frame(matriz.interaccion.adj[w,])
colnames(df.interaccion) <- colnames(matriz.interaccion.adj)
rownames(df.interaccion) <- rownames(matriz.interaccion.adj)[w]

mi_interaccion <-Reduce(intersect,list(rownames(df.grupo),rownames(df.obesos),rownames(df.interaccion)))

```

```{r}
## primera aproximacion
primera_aprox <-c("Haemophilus",
"Intestinibacter",
"Anaerotruncus",                          
"Christensenella",               
"Eisenbergiella",                        
"Granulicatella",                          
"Neisseria",                       
"Romboutsia",                          
"Saccharibacteria_genera_incertae_sedis",
"Staphylococcus")

primera_aprox[primera_aprox %in% mi_interaccion]
```


```{r}
library(ggstatsplot)

lista.graphos1 <- vector("list",length(primera_aprox))

for(i in 1:length(primera_aprox)){
  
  datos_i<-data.frame(variable=datos[,primera_aprox[i]])
  datos_i$interaccion<-interaccion
  levels(datos_i$interaccion) <- c("No.Obese_Female","Obese_Female",
                                   "No.Obese_PCOS","Obese_PCOS",
                                   "No.Obese_Male","Obese_Male")
  colnames(datos_i) <- c("variable","interaccion")
  lista.graphos1[[i]]<- ggbetweenstats(datos_i,x = interaccion,y=variable,type = "np",)+ylab(mi_interaccion[i])
  
}
  

```

```{r}
lista.graphos1
```


## SI  ahora agrupamos los subgrupos que son combinaciones de variables


```{r}
columnData.interaccion <- data.frame(obesidad=obesidad,grupo=grupo,interaccion=interaccion)
rownames(columnData.interaccion)<- pacientes
dds.interaccion <- DESeqDataSetFromMatrix(t(genus.abs),colData = columnData.interaccion,design = ~interaccion)
datos = as.data.frame(t(apply(assays(dds.interaccion.nb.2)[["mu"]], 2,log2)))

dds.interaccion <- DESeq2::estimateSizeFactors(dds.interaccion)
dds.interaccion <- DESeq2::estimateDispersions(dds.interaccion)

dds.interaccion.nb<- nbinomWaldTest(dds.interaccion)
dds.interaccion.lrt <- nbinomLRT(dds.interaccion,reduced = ~1)
# datos = as.data.frame(log2(assays(dds.interaccion)[["mu"]]/sizeFactors(dds.interaccion)))
csa <-getVarianceStabilizedData(dds.interaccion.nb)

datos <- as.data.frame(t(csa))
p1<-fn.densidad(datos = datos,fctr = interaccion)
p2 <- fn.boxplot(datos = datos,fctr = interaccion)
pca.log <- prcomp(t(datos),center = T,scale. = T)
pca.var <- 100*(pca.log$sdev^2/sum(pca.log$sdev^2))
p3<-graphPCA.methods(pca.log$rotation[,1:2],
                     datos.common = datos.common,
                     varianzas=pca.var)+ggtitle("PCA")
pca.spearman <- princomp(cor(t(datos),method = "spearman"),fix_sign = T,cor=T)

pca.spearman.scores <- pca.spearman$scores
pca.spearman.var <- 100*(pca.spearman$sdev^2/sum(pca.spearman$sdev^2))
p4<-graphPCA.methods(pca.spearman.scores[,1:2],
                 datos.common = datos.common,
                 varianzas = pca.spearman.var)+ggtitle("PCA spearman")



mds <- cmdscale(vegdist(datos,method = "euclidean"),eig = T,add = F,k=10)
varianzas <- 100*mds$eig/sum(mds$eig)
p5<-graphPCA.methods(datos=mds$points,
                     datos.common = datos.common,
                     varianzas = varianzas)+ggtitle("PCoA")


p5<-graphPCA.methods(datos=mds$points,
                     datos.common = datos.common,
                     varianzas = varianzas)+ggtitle("PCoA")



p11<-ggarrange(p1,p2,p3,p4,common.legend = T)
annotate_figure(p11,bottom = "media ajusta y rlog")
p5


```


#### univariante

```{r}
res <- results(dds.interaccion.nb)
res[which(res$padj<0.05),]
```


```{r}
res <- results(dds.interaccion.lrt)
res[which(res$padj<0.05),]
```

```{r}
res <- results(dds.interaccion.nb,contrast=c("interaccion","No.Obese_Female","Obese_Female"))
res[which(res$padj<0.05),]
```

```{r}
res <- results(dds.interaccion.nb,contrast=c("interaccion","No.Obese_Male","Obese_Male"))
res[which(res$padj<0.05),]
```

```{r}
res <- results(dds.interaccion.nb,contrast=c("interaccion","No.Obese_PCOS","Obese_PCOS"))
res[which(res$padj<0.05),]
```

```{r}
res <- results(dds.interaccion.nb,contrast=c("interaccion","No.Obese_Female","No.Obese_PCOS"))
res[which(res$padj<0.05),]
```

## Multivariante

```{r}
datos$grupo <- grupo
datos$obesidad <- obesidad
datos$interaccion <- interaccion
#### gruposde interes
females <- datos[datos$grupo=="Female",-c(95,96,97)]
males <- datos[datos$grupo=="Male",-c(95,96,97)]
pcos <- datos[datos$grupo=="PCOS",-c(95,96,97)]
obesos <- datos[datos$obesidad=="Obese",-c(95,96,97)]
No.Obesos <- datos[datos$obesidad=="No.Obese",-c(95,96,97)]
#### interacciones
females.no.obese <- datos[datos$interaccion=="No.Obese_Female",-c(95,96,97)]
females.obese <- datos[datos$interaccion=="Obese_Female",-c(95,96,97)]
males.no.obese <-  datos[datos$interaccion=="No.Obese_Male",-c(95,96,97)]
males.obese <-  datos[datos$interaccion=="Obese_Male",-c(95,96,97)]
pcos.no.obese <- datos[datos$interaccion=="No.Obese_PCOS",-c(95,96,97)]
pcos.obese <- datos[datos$interaccion=="Obese_PCOS",-c(95,96,97)]

```


```{r}
(res<-adonis2(datos[,-c(95,96,97)]~datos$interaccion))
```

```{r}
(res<-hotelling.test(females.no.obese,females.obese,shrinkage = T,var.equal = T,perm = T,B=100))
(res<-hotelling.test(males.no.obese,males.obese,shrinkage = T,var.equal = T,perm = T,B=100))
(res<-hotelling.test(pcos.no.obese,pcos.obese,shrinkage = T,var.equal = T,perm = T,B=100))
```

```{r}
(combinaciones.interaccion <- matrix(c(levels(interaccion)[1:3],levels(interaccion)[4:6]),ncol=2,nrow = 3))

lista.sujetos <- list(No.Obese_Female=females.no.obese,
                      Obese_Female=females.obese,
                      No.Obese_PCOS=pcos.no.obese,
                      Obese_PCOS=pcos.obese,
                      No.Obese_Male=males.no.obese,
                      Obese_Male=males.obese)
## interaccion

matriz.interaccion <- as.data.frame(matrix(NA,ncol=3,nrow=94))
colnames(matriz.interaccion)<-apply(combinaciones.interaccion, 1, function(x) paste(x[1],x[2],sep="_vs_"))
rownames(matriz.interaccion) <-colnames(genus.abs)

## grupo
for(j in 1:94){
for(i in 1:dim(combinaciones.interaccion)[1]){
  x<-lista.sujetos[[combinaciones.interaccion[i,1]]][,j]
  y <- lista.sujetos[[combinaciones.interaccion[i,2]]][,j]
  matriz.interaccion[j,i]<-wilcox.test(x,y)$p.value

  
}
}

matriz.interaccion.adj <- as.data.frame(t(apply(matriz.interaccion, 1, function(x) p.adjust(x,method="BH"))))

w <- which(apply(matriz.interaccion.adj, 1, function(x) any(x<0.05)))
df.interaccion <-data.frame(matriz.interaccion.adj[w,])
colnames(df.interaccion) <- colnames(matriz.interaccion.adj)
rownames(df.interaccion) <- rownames(matriz.interaccion.adj)[w]
```

```{r}

lista.graphos1 <- vector("list",length(primera_aprox))

for(i in 1:length(primera_aprox)){
  
  datos_i<-data.frame(variable=datos[,primera_aprox[i]])
  datos_i$interaccion<-interaccion
  levels(datos_i$interaccion) <- c("No.Obese_Female","Obese_Female",
                                   "No.Obese_PCOS","Obese_PCOS",
                                   "No.Obese_Male","Obese_Male")
  colnames(datos_i) <- c("variable","interaccion")
  lista.graphos1[[i]]<- ggbetweenstats(datos_i,x = interaccion,y=variable,type = "np",)+ylab(mi_interaccion[i])
  
}
  

```

