---
title: Integration of omics and clinical data of the hormonal, metabolic, inflammatory,
  and oxidative response in the different macronutrients of the diet
author: "Edmond Geraud"
subtitle: '`r params$subtitulo`'
date: '`r format(Sys.Date(),"%e de %B, %Y")`' 
output:
  html_document:
    toc: true
    toc_depth: 2
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    toc: true
header-includes:
  - \usepackage[english]{babel}
params:
  
  file_IP: Integromics_IPmarkers.xlsx
  file_metabolome: Integromics_Metabolome.xlsx
  file_general: Integromics_1.xlsx
  file_microbiota : integromics_microbiota.xlsx
  folder.data: ../../datos
  subtitulo: Inference without contrasts
geometry: margin=2cm
---
```{r setup, include=FALSE}
require(knitr)
# include this code chunk as-is to set options
opts_chunk$set(comment = NA, prompt = TRUE, tidy = T, 
               fig.width = 7, fig.height = 7,echo = TRUE, 
               message = FALSE, warning = FALSE, cache=TRUE, fig_caption = TRUE)
Sys.setlocale("LC_TIME", "C")
```

## Load the libraries



```{r libraries}

list.of.packages <- c("xlsx","kableExtra","dplyr","ggplot2","egg","cowplot","patchwork","gridExtra","UsingR","car","lattice","ggpubr","ggbreak","GGally","reshape2","ggcorrplot","corrplot","ggrepel","factoextra","chemometrics","sparsepca","vegan","ca","ade4","pls","OmicsPLS")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])] ## check for packages not installed 
if(length(new.packages)) install.packages(new.packages) ## install packages if necessary
res<-unlist(lapply(list.of.packages, require,character.only = T)) ## load packages needed for the session
# if(any(res==F)){
  list.of.packages[which(res==F)]  ## show those package if you have troubles to install.

list.of.bioc.packages <- c("phyloseq","pcaMethods","ropls","mixOmics","limma","DESeq2","edgeR")
new.packages.bioc <- list.of.packages[!(list.of.bioc.packages %in% installed.packages()[,"Package"])]
if(length(new.packages.bioc)) BiocManager::install(new.packages.bioc)
res.bioc <- unlist(lapply(list.of.bioc.packages, require,character.only = T)) 
if(any(res.bioc==F)){
  list.of.packages[which(res.bioc==F)]  ## show those package if you have troubles to install.
}
```

# Funciones
```{r functions}


proces_bacteria <- function(datos,tipo){

  Order <- bacteria_phylum.abs$Order
  variables_in_bacteria <- general_data[general_data$Orden %in% Order , ]
  Sample<-variables_in_bacteria$Paciente


  if(tipo=="genera"){

    Order <- Order

    X <- datos[-1,-1]
    X<-apply(X,2,as.numeric)
    colnames(X) <- datos[1,2:ncol(datos)]
    rownames(X) <- Sample
    X.rel <- 100*X/rowSums(X)


  } else if(tipo=="phylum"){

    X <- datos[,-c(1:2)]
    colnames(X) <- colnames(datos)[3:ncol(datos)]
    rownames(X) <- Sample
    X.rel<-100*X/rowSums(X)
  }

  GROUP <-variables_in_bacteria$GROUP
  SEX <- variables_in_bacteria$SEX
  OBESE<- variables_in_bacteria$OBESE

  X<-as.data.frame(X)
  X.rel<-as.data.frame(X.rel)

  X$GROUP<-GROUP
  X$SEX<-SEX
  X$OBESE<-OBESE


  X.rel$GROUP<-GROUP
  X.rel$SEX<-SEX
  X.rel$OBESE<-OBESE

  return(list(abs=X,rel=X.rel))

}


low.count.removal = function(
                        data, # OTU count data frame of size n (sample) x p (OTU)
                        percent=0.01 # cutoff chosen
                        ){
    keep.otu = which(colSums(data)*100/(sum(colSums(data))) > percent)
    data.filter = data[,keep.otu]
    return(list(data.filter = data.filter, keep.otu = keep.otu))
}

low.count.removal.counts = function(
                        data, # OTU count data frame of size n (sample) x p (OTU)
                        abundance=10 # cutoff chosen
                        ){
    keep.otu <- which(colSums(data)>abundance)
    data.filter <-data[,keep.otu]
    return(list(data.filter = data.filter, keep.otu = keep.otu))
}

relative.fn <-function(data){
  
  return(100*data/rowSums(data))
  
}


common.bacteria <- function(datos_i,datos.comunes){
  

  ## retorna los indices con las variables comunes

  interaccion <- interaction(datos.comunes$GROUP,datos.comunes$OBESE)
  datos.comunes$interaccion <- interaccion
  grupo <-levels(interaccion)
  w<-vector(mode="list",length=6)
  names(w)<-levels(interaccion)
  for(i in 1:6){
   w[[names(w)[i]]]<-which(apply(as.matrix(datos_i[
    rownames(datos_i)%in%
      datos.comunes[datos.comunes$interaccion==grupo[i],"Paciente"],]),2,function(x) sum(x)==0))
  }

  w2<-vector(mode="list",length=6)
  names(w2)<-levels(interaccion)
  for(i in 1:6){
   w2[[names(w2)[i]]]<-which(apply(as.matrix(datos_i[
    rownames(datos_i)%in%
      datos.comunes[datos.comunes$interaccion==grupo[i],"Paciente"],]),2,function(x) sum(x)>0))
    }

  w3<-lapply(w2, names)
  common_variables <-Reduce(intersect,x = w3)

  return(common_variables)
}

fn.densidad <- function(datos,fctr){
  column_var=c("interaccion","variable","value")
  columnas <- colnames(datos)
  filas <- rownames(datos)
  X<-data.frame(apply(datos, 2, as.numeric))
  colnames(X)<-columnas
  rownames(X)<-filas
  X$fctr<-fctr
  X.melt <- melt(X)
  colnames(X.melt)<-column_var
  p<-ggplot(data=X.melt,aes(x=value,color=interaccion))+geom_density()
  
  return(p)
}

fn.boxplot<-function(datos,fctr){
  column_var=c("interaccion","variable","value")
  columnas <- colnames(datos)
  filas <- rownames(datos)
  X<-data.frame(apply(datos, 2, as.numeric))
  colnames(X)<-columnas
  rownames(X)<-filas
  X$fctr<-fctr
  X.melt <- melt(X)
  colnames(X.melt)<-column_var
  
  p<-ggplot(X.melt,aes(x=interaccion,y=value,fill=interaccion))+geom_boxplot()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  return(p)
}


graphPCA.methods <- function(datos,datos.common,varianzas){
  
  
  sujetos <-as.data.frame(datos)
  sexo<-datos.common$GROUP
  obesidad<-datos.common$OBESE
  colnames(sujetos)<-c("PC1","PC2")
  p1<-ggplot(sujetos,aes(x=PC1, y=PC2)) +
     geom_hline(yintercept = 0, color = "gray70") +
     geom_vline(xintercept = 0, color = "gray70") +
     geom_point(aes(color = sexo,shape=obesidad,fill=obesidad), alpha = 0.55, size = 3) +
     coord_cartesian(xlim = c(min(sujetos[,1]),max(sujetos[,1]))) +
     scale_color_manual(values=c("black","red","green"))+scale_shape_manual(values=c(17,15))+xlab(paste("PC1",round(varianzas[1],2),"%"))+ylab(paste("PC2",round(varianzas[2],2),"%"))
  return(p1)
}

```


# Cargamos los datos

## Datos clinicos

```{r}
general_data <- read.xlsx(file.path(params$folder.data,params$file_general),sheetIndex =1)
general_data$SEX <- factor(general_data$SEX, levels = c(0, 2), labels = c("Female", "Male"))
general_data$OBESE <- factor(general_data$OBESE, levels = c(0, 1), labels = c("No Obese", "Obese"))
general_data$GROUP <- factor(general_data$GROUP, levels = c(0, 1, 2), labels = c("Female",  "PCOS", "Male"))


```


## Datos IP

```{r}

IP_file<-file.path(params$folder.data,file=params$file_IP) ## file path of the data
IP.tmp<-read.xlsx(IP_file,sheetIndex = 1)
IP<-IP.tmp[,-7] # remove the missin values column
IP$SEX <- factor(IP$SEX,levels=c(0,2),labels = c("Female","Male"))
IP$OBESE <- factor(IP$OBESE,levels=c(0,1),labels = c("No Obese","Obese"))
IP$GROUP <- factor(IP$GROUP,levels=c(0,1,2),labels = c("Female","PCOS","Male"))

```

## Datos Metaboloma

```{r}
metabolome_file<-file.path(params$folder.data,file=params$file_metabolome) ## file path of the data
metabolome.tmp<-read.xlsx(metabolome_file,sheetIndex = 1) # read the data
metabolites_name <- read.xlsx(metabolome_file,sheetIndex = 2)$Metabolitos # read the names of the metabolites
any(is.na(metabolome.tmp)) # there is a row of missing values.
metabolome<-metabolome.tmp[-nrow(metabolome.tmp),] ## remove the missing value row
metabolome$GROUP<-general_data$GROUP ## add GROUP variable 
metabolome$OBESE <- general_data$OBESE ## add OBESE variables.

```


## Datos microbioma

```{r}
bacteria_phylum.abs <- as.data.frame(read.xlsx(file.path(params$folder.data,params$file_microbiota), sheetIndex = 1))
bacteria_genera.abs <- as.data.frame(t(read.xlsx(file.path(params$folder.data,params$file_microbiota), sheetIndex = 3)))
phylum.list <- proces_bacteria(bacteria_phylum.abs,tipo="phylum")
genera.list <- proces_bacteria(bacteria_genera.abs,tipo="genera")

phylum.abs <- phylum.list$abs;phylum.rel<-phylum.list$rel
genera.abs <- genera.list$abs;genera.rel<-genera.list$rel

```

# Preparando los datos para la integracion
```{r}
variables_in_bacteria <- general_data[general_data$Orden %in% bacteria_phylum.abs$Order, ]

```

## Datos Clinicos
**NOTE we are going to include also on both levels of integration body measures**
We exclude ISI because it is measured after the load of glucose, the mean values, to exclude redundancy of the data, ant the postprandial levels. 


```{r}
sujetos <- variables_in_bacteria$Paciente
Clin.tmp <- variables_in_bacteria[,-c(1,2,3,4,5)] # get rid off orden, paciente, sex, group and obese
auc_clin <- colnames(Clin.tmp)[grep("AUC",colnames(Clin.tmp))]
mean_basal_clin<-colnames(Clin.tmp)[grep("_B$",colnames(Clin.tmp))]
sog <- colnames(Clin.tmp)[grep("^SO[GLP]",colnames(Clin.tmp))]
variables_clin <- colnames(Clin.tmp)
testosterona <-colnames(Clin.tmp)[grep("TEST",colnames(Clin.tmp))]
interr<-colnames(Clin.tmp)[grep("interaccion",colnames(Clin.tmp))]
ratio <- colnames(Clin.tmp)[grep("Ratio",colnames(Clin.tmp))]
w<-c(which(variables_clin %in% auc_clin),which(variables_clin %in% mean_basal_clin),which(variables_clin %in% sog),which(variables_clin %in% testosterona),
which(variables_clin %in% interr),which(variables_clin %in% ratio))
Clin <- Clin.tmp[,-w] 
rownames(Clin) <- sujetos
colnames(Clin)
```


## Datos IP
```{r}
IP.tmp <- IP[IP$Paciente %in% variables_in_bacteria$Paciente,]
redundant_variables <- c("Orden","Paciente","SEX","GROUP","OBESE","BMI")
variables_ip <- colnames(IP.tmp)
auc_ip <- variables_ip[grep("AUC",variables_ip)]
mean_ip <- variables_ip[grep("_0$",variables_ip)]
w <- c(which(variables_ip %in% redundant_variables),which(variables_ip %in% auc_ip),which(variables_ip %in% mean_ip))
IP.basal <- IP.tmp[,-w]
rownames(IP.basal) <- sujetos

```

## Datos metaboloma

```{r}
metabolome.basal.tmp <- metabolome[metabolome$Order %in% variables_in_bacteria$Orden,]
redundant_data_meta <- c("Order","Sample","GROUP","OBESE")
variables_meta <- colnames(metabolome.basal.tmp)
mean_meta <- variables_meta[grep("_[GLP]0$",variables_meta)]
auc_meta <- variables_meta[grep("AU[CG]",variables_meta)]
w <- c(which(variables_meta %in% redundant_data_meta),which(variables_meta %in% mean_meta),which(variables_meta %in% auc_meta))
metabolome.basal <- metabolome.basal.tmp[,-w]
rownames(metabolome.basal) <- sujetos
colnames(metabolome.basal)

```
# Datos que necesitamos

```{r}
pacientes <- variables_in_bacteria$Paciente
```

## Datos microbioma

### Datos genero

```{r}

genera.abs.tmp <- genera.abs
redundant_data_ph <- c("GROUP","SEX","OBESE")
genus.tmp.tmp <- genera.abs.tmp[,-c(ncol(genera.abs.tmp)-2,ncol(genera.abs.tmp)-1,ncol(genera.abs.tmp))]
genus <- genus.tmp.tmp[,colSums(genus.tmp.tmp)>0]
rownames(genus) <- sujetos
## removal of genus unidentified

genus<-genus[,-grep("^ud-",colnames(genus))]

common_genus <-common.bacteria(genus,variables_in_bacteria)

genus<-genus[,common_genus]
```

# Preparando inferencia

Como los disenyos recomiendan hacer la interaccion de los factores y luego hacer contrastes (ver documentacion)
"
Initial note: Many users begin to add interaction terms to the design formula, when in fact a much simpler approach would give all the results tables that are desired. We will explain this approach first, because it is much simpler to perform. If the comparisons of interest are, for example, the effect of a condition for different sets of samples, a simpler approach than adding interaction terms explicitly to the design formula is to perform the following steps:"

## Primero visualizamos con graficos de barras via phyloseq

Pero antes removemos aquellas que tienen menos de 10 abundancia

```{r}
genus.abs <- low.count.removal.counts(genus,10)$data.filter

```



# Preparando para la inferencia GENERO

Primero vamos a filtrar las variables comunes por aquellas que tengan una minima abundancia de 10
Preparamos el enteorno phyloseq por si las moscas

```{r}
grupo<-variables_in_bacteria$GROUP
obesos <- factor(gsub(" ",".",variables_in_bacteria$OBESE))
interaccion<-factor(paste(obesos,grupo,sep="_"))
otumat<-t(genus.abs)
rownames(otumat)<-paste0("OTU",1:nrow(otumat))
colnames(otumat)<-pacientes
taxmat <- cbind(colnames(genus.abs),interaccion)
colnames(taxmat) <- c("Genus","interaccion")
rownames(taxmat)<-rownames(otumat)
OTU <- otu_table(otumat,taxa_are_rows = T)
TAX <- tax_table(taxmat)
physeq <- phyloseq(OTU,TAX)
```

### Vemos los datos crudos
```{r}
genus.premelt <- genus.abs
genus.premelt$interaccion<-interaccion
genus.melt <- melt(genus.premelt)
ggplot(genus.melt,aes(x=interaccion,y=value))+geom_bar(aes(color=variable,fill=variable),stat="identity",position = position_dodge(0.8))+theme(legend.position="none")
```
## Transformaciones
```{r}
datos.common <- variables_in_bacteria[,c("GROUP","OBESE","Paciente")]
```

### log2(x+1)


```{r}
p1<-fn.densidad(datos = log2(genus.abs+1),fctr = interaccion)
p2 <- fn.boxplot(datos = log2(genus.abs+1),fctr = interaccion)
pca.log <- prcomp(t(log2(genus.abs+1)),center = F,scale. = F)
pca.var <- 100*(pca.log$sdev^2/sum(pca.log$sdev^2))
p3<-graphPCA.methods(pca.log$rotation[,1:2],
                     datos.common = datos.common,
                     varianzas=pca.var)+ggtitle("PCA")
plsda.log <- plsda(log2(genus.abs+1),Y=interaccion,scale = T)
plsda.var <-100*plsda.log$prop_expl_var$X
p4<-graphPCA.methods(datos=plsda.log$variates$X,
                     datos.common = datos.common,
                     varianzas = plsda.var)+ggtitle("PLS-DA")
p5<-ggarrange(p1,p2,common.legend = T)
annotate_figure(p5,bottom = "log2(x+1)")
p6<-ggarrange(p3,p4,ncol = 2,common.legend = T)
annotate_figure(p6,bottom = "log2(x+1)")

```



### transformacion (segun mixomics)
```{r}
lib.size <- apply(genus.abs, 1, sum)
barplot(lib.size,las=2,names.arg = interaccion,cex.names = 0.5,col=interaccion)
```

#### CLR and ILR
```{r}

data.clr <- logratio.transfo(as.matrix(genus.abs), 
                             logratio = 'CLR', offset = 1) 

data.ilr <- logratio.transfo(as.matrix(genus.abs),logratio = "ILR",offset = 1)
```


##### CLR

```{r}
p1<-fn.densidad(datos = data.clr,fctr = interaccion)
p2 <- fn.boxplot(datos = data.clr,fctr = interaccion)
pca.log <- prcomp(t(data.clr),center = T,scale. = T)
pca.var <- 100*(pca.log$sdev^2/sum(pca.log$sdev^2))
p3<-graphPCA.methods(pca.log$rotation[,1:2],
                     datos.common = datos.common,
                     varianzas=pca.var)+ggtitle("PCA")
pca.spearman <- princomp(cor(t(data.clr),method = "spearman"),fix_sign = T,cor=T)

pca.spearman.scores <- pca.spearman$scores
pca.spearman.var <- 100*(pca.spearman$sdev^2/sum(pca.spearman$sdev^2))
p4<-graphPCA.methods(pca.spearman.scores[,1:2],
                 datos.common = datos.common,
                 varianzas = pca.spearman.var)+ggtitle("PCA spearman")



plsda.log <- plsda(data.clr,Y=interaccion,scale = F)
plsda.var <-100*plsda.log$prop_expl_var$X
graphPCA.methods(datos=plsda.log$variates$X,
                     datos.common = datos.common,
                     varianzas = plsda.var)+ggtitle("PLS-DA CLR")


p7<-ggarrange(p1,p2,common.legend = T)
annotate_figure(p7,bottom = "CLR")
p8<-ggarrange(p3,p4,ncol = 2,common.legend = T)
annotate_figure(p8,bottom = "CLR")

```

##### ILR

```{r}
p1<-fn.densidad(datos = data.ilr,fctr = interaccion)
p2 <- fn.boxplot(datos = data.ilr,fctr = interaccion)
pca.log <- prcomp(t(data.ilr),center = T,scale. = T)
pca.var <- 100*(pca.log$sdev^2/sum(pca.log$sdev^2))
p3<-graphPCA.methods(pca.log$rotation[,1:2],
                     datos.common = datos.common,
                     varianzas=pca.var)+ggtitle("PCA")
pca.spearman <- princomp(cor(t(data.ilr),method = "spearman"),fix_sign = T,cor=T)

pca.spearman.scores <- pca.spearman$scores
pca.spearman.var <- 100*(pca.spearman$sdev^2/sum(pca.spearman$sdev^2))
p4<-graphPCA.methods(pca.spearman.scores[,1:2],
                 datos.common = datos.common,
                 varianzas = pca.spearman.var)+ggtitle("PCA spearman")



plsda.log <- plsda(data.ilr,Y=interaccion,scale = F)
plsda.var <-100*plsda.log$prop_expl_var$X
graphPCA.methods(datos=plsda.log$variates$X,
                     datos.common = datos.common,
                     varianzas = plsda.var)+ggtitle("PLS-DA ILR")


p7<-ggarrange(p1,p2,common.legend = T)
annotate_figure(p7,bottom = "ILR")
p8<-ggarrange(p3,p4,ncol = 2,common.legend = T)
annotate_figure(p8,bottom = "ILR")

```



### Ahora vamos a probar con limma edgeR sin CLR

#### TMM
```{r}
columnData <- data.frame(interaccion=interaccion)
rownames(columnData) <- pacientes
design <- model.matrix(~0+interaccion)
dge <- DGEList(counts=t(genus.abs),samples=columnData,group = interaccion)
nf <-calcNormFactors(dge,refColumn = 1,method="TMM")
voom.norm.tmm <- voom(nf,plot = T)
datos<-t(voom.norm.tmm$E)
p1<-fn.densidad(datos = datos,fctr = interaccion)
p2 <- fn.boxplot(datos = datos,fctr = interaccion)
pca.log <- prcomp(t(datos),center = T,scale. = T)
pca.var <- 100*(pca.log$sdev^2/sum(pca.log$sdev^2))
p3<-graphPCA.methods(pca.log$rotation[,1:2],
                     datos.common = datos.common,
                     varianzas=pca.var)+ggtitle("PCA")
pca.spearman <- princomp(cor(t(datos),method = "spearman"),fix_sign = T,cor=T)

pca.spearman.scores <- pca.spearman$scores
pca.spearman.var <- 100*(pca.spearman$sdev^2/sum(pca.spearman$sdev^2))
p4<-graphPCA.methods(pca.spearman.scores[,1:2],
                 datos.common = datos.common,
                 varianzas = pca.spearman.var)+ggtitle("PCA spearman")



plsda.log <- plsda(datos,Y=interaccion,scale = F)
plsda.var <-100*plsda.log$prop_expl_var$X
p5<-graphPCA.methods(datos=plsda.log$variates$X,
                     datos.common = datos.common,
                     varianzas = plsda.var)+ggtitle("PLS-DA")



plsda.log.clr <- plsda(datos,Y=interaccion,scale = F,logratio = "CLR")
plsda.log.clr.var <-100*plsda.log.clr$prop_expl_var$X

p6<-graphPCA.methods(datos=plsda.log.clr$variates$X,
                     datos.common = datos.common,
                     varianzas = plsda.log.clr.var)+ggtitle("PLS-DA CLR")


p7<-ggarrange(p1,p2,common.legend = T)
annotate_figure(p7,bottom = "TMM")
p8<-ggarrange(p3,p4,ncol = 2,common.legend = T)
annotate_figure(p8,bottom = "TMM")
p9<-ggarrange(p5,p6,ncol = 2,common.legend = T)
annotate_figure(p9,bottom = "TMM")




```



#### TMMwsp
```{r}
columnData <- data.frame(interaccion=interaccion)
rownames(columnData) <- pacientes
design <- model.matrix(~0+interaccion)
dge <- DGEList(counts=t(genus.abs),samples=columnData,group = interaccion)
nf <-calcNormFactors(dge,refColumn = 1,method="TMMwsp")
voom.norm.tmm <- voom(nf,plot = T)
datos<-t(voom.norm.tmm$E)
p1<-fn.densidad(datos = datos,fctr = interaccion)
p2 <- fn.boxplot(datos = datos,fctr = interaccion)
pca.log <- prcomp(t(datos),center = T,scale. = T)
pca.var <- 100*(pca.log$sdev^2/sum(pca.log$sdev^2))
p3<-graphPCA.methods(pca.log$rotation[,1:2],
                     datos.common = datos.common,
                     varianzas=pca.var)+ggtitle("PCA")
pca.spearman <- princomp(cor(t(datos),method = "spearman"),fix_sign = T,cor=T)

pca.spearman.scores <- pca.spearman$scores
pca.spearman.var <- 100*(pca.spearman$sdev^2/sum(pca.spearman$sdev^2))
p4<-graphPCA.methods(pca.spearman.scores[,1:2],
                 datos.common = datos.common,
                 varianzas = pca.spearman.var)+ggtitle("PCA spearman")



plsda.log <- plsda(datos,Y=interaccion,scale = F)
plsda.var <-100*plsda.log$prop_expl_var$X
p5<-graphPCA.methods(datos=plsda.log$variates$X,
                     datos.common = datos.common,
                     varianzas = plsda.var)+ggtitle("PLS-DA")



plsda.log.clr <- plsda(datos,Y=interaccion,scale = F,logratio = "CLR")
plsda.log.clr.var <-100*plsda.log.clr$prop_expl_var$X

p6<-graphPCA.methods(datos=plsda.log.clr$variates$X,
                     datos.common = datos.common,
                     varianzas = plsda.log.clr.var)+ggtitle("PLS-DA CLR")


p7<-ggarrange(p1,p2,common.legend = T)
annotate_figure(p7,bottom = "TMMwsp")
p8<-ggarrange(p3,p4,ncol = 2,common.legend = T)
annotate_figure(p8,bottom = "TMMwsp")
p9<-ggarrange(p5,p6,ncol = 2,common.legend = T)
annotate_figure(p9,bottom = "TMMwsp")




```


#### RLE
```{r}
columnData <- data.frame(interaccion=interaccion)
rownames(columnData) <- pacientes
design <- model.matrix(~0+interaccion)
dge <- DGEList(counts=t(genus.abs),samples=columnData,group = interaccion)
nf <-calcNormFactors(dge,refColumn = NULL,method="RLE")
voom.norm.tmm <- voom(nf,plot = T)
datos<-t(voom.norm.tmm$E)
p1<-fn.densidad(datos = datos,fctr = interaccion)
p2 <- fn.boxplot(datos = datos,fctr = interaccion)
pca.log <- prcomp(t(datos),center = T,scale. = T)
pca.var <- 100*(pca.log$sdev^2/sum(pca.log$sdev^2))
p3<-graphPCA.methods(pca.log$rotation[,1:2],
                     datos.common = datos.common,
                     varianzas=pca.var)+ggtitle("PCA")
pca.spearman <- princomp(cor(t(datos),method = "spearman"),fix_sign = T,cor=T)

pca.spearman.scores <- pca.spearman$scores
pca.spearman.var <- 100*(pca.spearman$sdev^2/sum(pca.spearman$sdev^2))
p4<-graphPCA.methods(pca.spearman.scores[,1:2],
                 datos.common = datos.common,
                 varianzas = pca.spearman.var)+ggtitle("PCA spearman")



plsda.log <- plsda(datos,Y=interaccion,scale = F)
plsda.var <-100*plsda.log$prop_expl_var$X
p5<-graphPCA.methods(datos=plsda.log$variates$X,
                     datos.common = datos.common,
                     varianzas = plsda.var)+ggtitle("PLS-DA")



plsda.log.clr <- plsda(datos,Y=interaccion,scale = F,logratio = "CLR")
plsda.log.clr.var <-100*plsda.log.clr$prop_expl_var$X

p6<-graphPCA.methods(datos=plsda.log.clr$variates$X,
                     datos.common = datos.common,
                     varianzas = plsda.log.clr.var)+ggtitle("PLS-DA CLR")


p7<-ggarrange(p1,p2,common.legend = T)
annotate_figure(p7,bottom = "RLE")
p8<-ggarrange(p3,p4,ncol = 2,common.legend = T)
annotate_figure(p8,bottom = "RLE")
p9<-ggarrange(p5,p6,ncol = 2,common.legend = T)
annotate_figure(p9,bottom = "RLE")




```


#### upperquartile
```{r}
columnData <- data.frame(interaccion=interaccion)
rownames(columnData) <- pacientes
design <- model.matrix(~0+interaccion)
dge <- DGEList(counts=t(genus.abs),samples=columnData,group = interaccion)
nf <-calcNormFactors(dge,refColumn = NULL,method="upperquartile")
voom.norm.tmm <- voom(nf,plot = T)
datos<-t(voom.norm.tmm$E)
p1<-fn.densidad(datos = datos,fctr = interaccion)
p2 <- fn.boxplot(datos = datos,fctr = interaccion)
pca.log <- prcomp(t(datos),center = T,scale. = T)
pca.var <- 100*(pca.log$sdev^2/sum(pca.log$sdev^2))
p3<-graphPCA.methods(pca.log$rotation[,1:2],
                     datos.common = datos.common,
                     varianzas=pca.var)+ggtitle("PCA")
pca.spearman <- princomp(cor(t(datos),method = "spearman"),fix_sign = T,cor=T)

pca.spearman.scores <- pca.spearman$scores
pca.spearman.var <- 100*(pca.spearman$sdev^2/sum(pca.spearman$sdev^2))
p4<-graphPCA.methods(pca.spearman.scores[,1:2],
                 datos.common = datos.common,
                 varianzas = pca.spearman.var)+ggtitle("PCA spearman")



plsda.log <- plsda(datos,Y=interaccion,scale = F)
plsda.var <-100*plsda.log$prop_expl_var$X
p5<-graphPCA.methods(datos=plsda.log$variates$X,
                     datos.common = datos.common,
                     varianzas = plsda.var)+ggtitle("PLS-DA")



plsda.log.clr <- plsda(datos,Y=interaccion,scale = F,logratio = "CLR")
plsda.log.clr.var <-100*plsda.log.clr$prop_expl_var$X

p6<-graphPCA.methods(datos=plsda.log.clr$variates$X,
                     datos.common = datos.common,
                     varianzas = plsda.log.clr.var)+ggtitle("PLS-DA CLR")


p7<-ggarrange(p1,p2,common.legend = T)
annotate_figure(p7,bottom = "upperquartile")
p8<-ggarrange(p3,p4,ncol = 2,common.legend = T)
annotate_figure(p8,bottom = "upperquartile")
p9<-ggarrange(p5,p6,ncol = 2,common.legend = T)
annotate_figure(p9,bottom = "upperquartile")




```





#### none
```{r}
columnData <- data.frame(interaccion=interaccion)
rownames(columnData) <- pacientes
design <- model.matrix(~0+interaccion)
dge <- DGEList(counts=t(genus.abs),samples=columnData,group = interaccion)
nf <-calcNormFactors(dge,refColumn = NULL,method="none")
voom.norm.tmm <- voom(nf,plot = T)
datos<-t(voom.norm.tmm$E)
p1<-fn.densidad(datos = datos,fctr = interaccion)
p2 <- fn.boxplot(datos = datos,fctr = interaccion)
pca.log <- prcomp(t(datos),center = T,scale. = T)
pca.var <- 100*(pca.log$sdev^2/sum(pca.log$sdev^2))
p3<-graphPCA.methods(pca.log$rotation[,1:2],
                     datos.common = datos.common,
                     varianzas=pca.var)+ggtitle("PCA")
pca.spearman <- princomp(cor(t(datos),method = "spearman"),fix_sign = T,cor=T)

pca.spearman.scores <- pca.spearman$scores
pca.spearman.var <- 100*(pca.spearman$sdev^2/sum(pca.spearman$sdev^2))
p4<-graphPCA.methods(pca.spearman.scores[,1:2],
                 datos.common = datos.common,
                 varianzas = pca.spearman.var)+ggtitle("PCA spearman")



plsda.log <- plsda(datos,Y=interaccion,scale = F)
plsda.var <-100*plsda.log$prop_expl_var$X
p5<-graphPCA.methods(datos=plsda.log$variates$X,
                     datos.common = datos.common,
                     varianzas = plsda.var)+ggtitle("PLS-DA")



plsda.log.clr <- plsda(datos,Y=interaccion,scale = F,logratio = "CLR")
plsda.log.clr.var <-100*plsda.log.clr$prop_expl_var$X

p6<-graphPCA.methods(datos=plsda.log.clr$variates$X,
                     datos.common = datos.common,
                     varianzas = plsda.log.clr.var)+ggtitle("PLS-DA CLR")


p7<-ggarrange(p1,p2,common.legend = T)
annotate_figure(p7,bottom = "none")
p8<-ggarrange(p3,p4,ncol = 2,common.legend = T)
annotate_figure(p8,bottom = "none")
p9<-ggarrange(p5,p6,ncol = 2,common.legend = T)
annotate_figure(p9,bottom = "none")




```

# La mejor separacion del microbioma es con TMMwsp y CLR

```{r}
columnData <- data.frame(interaccion=interaccion)
rownames(columnData) <- pacientes
design <- model.matrix(~0+interaccion)
dge <- DGEList(counts=t(genus.abs),samples=columnData,group = interaccion)
nf <-calcNormFactors(dge,refColumn = 1,method="TMMwsp")
voom.norm.tmm <- voom(nf,plot = T)
datos<-t(voom.norm.tmm$E)
p1<-fn.densidad(datos = datos,fctr = interaccion)
p2 <- fn.boxplot(datos = datos,fctr = interaccion)
pca.log <- prcomp(t(datos),center = T,scale. = T)
pca.var <- 100*(pca.log$sdev^2/sum(pca.log$sdev^2))
p3<-graphPCA.methods(pca.log$rotation[,1:2],
                     datos.common = datos.common,
                     varianzas=pca.var)+ggtitle("PCA")
pca.spearman <- princomp(cor(t(datos),method = "spearman"),fix_sign = T,cor=T)

pca.spearman.scores <- pca.spearman$scores
pca.spearman.var <- 100*(pca.spearman$sdev^2/sum(pca.spearman$sdev^2))
p4<-graphPCA.methods(pca.spearman.scores[,1:2],
                 datos.common = datos.common,
                 varianzas = pca.spearman.var)+ggtitle("PCA spearman")



plsda.log <- plsda(datos,Y=interaccion,scale = F)
plsda.var <-100*plsda.log$prop_expl_var$X
p5<-graphPCA.methods(datos=plsda.log$variates$X,
                     datos.common = datos.common,
                     varianzas = plsda.var)+ggtitle("PLS-DA")



plsda.log.clr <- plsda(datos,Y=interaccion,scale = F,logratio = "CLR")
plsda.log.clr.var <-100*plsda.log.clr$prop_expl_var$X

p6<-graphPCA.methods(datos=plsda.log.clr$variates$X,
                     datos.common = datos.common,
                     varianzas = plsda.log.clr.var)+ggtitle("PLS-DA CLR")


p7<-ggarrange(p1,p2,common.legend = T)
annotate_figure(p7,bottom = "TMMwsp")
p8<-ggarrange(p3,p4,ncol = 2,common.legend = T)
annotate_figure(p8,bottom = "TMMwsp")
p9<-ggarrange(p5,p6,ncol = 2,common.legend = T)
annotate_figure(p9,bottom = "TMMwsp")




```

## Vamos con inferneicia univariante
```{r}
library(DEFormats)
dds <- as.DESeqDataSet(nf)
voom.norm.tmm<-voom(assay(dds),plot = F)
comparaciones<-makeContrasts(female_obese="No.Obese_Female-Obese_Female",
                 male_obese = "No.Obese_Male-Obese_Male",
                 pcos_obese = "No.Obese_PCOS-Obese_PCOS",
                 female_vs_pcos = "No.Obese_Female-No.Obese_PCOS",
                 male_vs_pcos = "No.Obese_Male-No.Obese_PCOS",
                 female_vs_male = "No.Obese_Female-No.Obese_Male",
                 female_vs_pcos.O = "Obese_Female-Obese_PCOS",
                 male_vs_pcos.O = "Obese_Male-Obese_PCOS",
                 female_vs_male.O = "Obese_Female-Obese_Male"
                 ,levels = design)

fit<-lmFit(voom.norm.tmm,design = voom.norm.tmm$design)
fit <- eBayes(fit)
summary(decideTests(fit))
```

```{r}
sse <- SummarizedExperiment(t(genus.abs),colData = columnData)
dds <- DESeqDataSet(sse,design = ~0+interaccion)
dds$condition <- interaccion

dds<-DESeq(dds)
```
```{r}
res<-results(dds)
```

```{r}
dds <- DESeqDataSetFromMatrix(t(genus.abs),colData = columnData,design = ~0+interaccion)
dds <- DESeq(dds)
```

```{r}
res.fem_obese <- results(dds,c("interaccion","No.Obese_Female","Obese_Female"))
res.fem_obese[order(res.fem_obese$padj),]
```
```{r}
res.male_obese <- results(dds,c("interaccion","No.Obese_Male","Obese_Male"))
res.male_obese[order(res.male_obese$padj),]
```

```{r}
res.pcos_obese <- results(dds,c("interaccion","No.Obese_PCOS","Obese_PCOS"))
res.pcos_obese[order(res.pcos_obese$padj),]
```

```{r}
res.fem_pcos <- results(dds,c("interaccion","No.Obese_Female","No.Obese_PCOS"))
res.fem_pcos[order(res.fem_pcos$padj),]
```

```{r}
res.fem_pcos.O <- results(dds,c("interaccion","Obese_Female","Obese_PCOS"))
res.fem_pcos.O[order(res.fem_pcos.O$padj),]
```

```{r}
res.male_pcos.O <- results(dds,c("interaccion","Obese_Male","Obese_PCOS"))
res.male_pcos.O[order(res.male_pcos.O$padj),]
```
```{r}
res.female_pcos.O <- results(dds,c("interaccion","No.Obese_Female","Obese_PCOS"))
res.female_pcos.O[order(res.female_pcos.O$padj),]
```

```{r}
res.female_pcos <- results(dds,c("interaccion","No.Obese_Female","No.Obese_PCOS"))
res.female_pcos[order(res.female_pcos$padj),]
```
```{r}
res[order(res$padj),]
```

```{r}
resultsNames(dds)
```


```{r}
X<-counts(dds,normalized=T)
datos<-t(X)
p1<-fn.densidad(datos = datos,fctr = interaccion)
p2 <- fn.boxplot(datos = datos,fctr = interaccion)
pca.log <- prcomp(t(datos),center = T,scale. = T)
pca.var <- 100*(pca.log$sdev^2/sum(pca.log$sdev^2))
p3<-graphPCA.methods(pca.log$rotation[,1:2],
                     datos.common = datos.common,
                     varianzas=pca.var)+ggtitle("PCA")
pca.spearman <- princomp(cor(t(datos),method = "spearman"),fix_sign = T,cor=T)

pca.spearman.scores <- pca.spearman$scores
pca.spearman.var <- 100*(pca.spearman$sdev^2/sum(pca.spearman$sdev^2))
p4<-graphPCA.methods(pca.spearman.scores[,1:2],
                 datos.common = datos.common,
                 varianzas = pca.spearman.var)+ggtitle("PCA spearman")



plsda.log <- plsda(datos,Y=interaccion,scale = F)
plsda.var <-100*plsda.log$prop_expl_var$X
p5<-graphPCA.methods(datos=plsda.log$variates$X,
                     datos.common = datos.common,
                     varianzas = plsda.var)+ggtitle("PLS-DA")




p7<-ggarrange(p1,p2,common.legend = T)
annotate_figure(p7,bottom = "TMMwsp")
p8<-ggarrange(p3,p4,ncol = 2,common.legend = T)
annotate_figure(p8,bottom = "TMMwsp")
p9<-ggarrange(p5,p6,ncol = 2,common.legend = T)
p5

```



# PRUEBA

```{r}
dds <- DESeqDataSetFromMatrix(t(genus.abs),colData = columnData,design = ~0+interaccion)
dds <- DESeq2::estimateSizeFactors(dds)
dds <- DESeq2::estimateDispersions(dds)
dds.nb <- nbinomWaldTest(dds)
mu<-assays(dds.nb)$mu

datos<-(t(log2(mu)))
p1<-fn.densidad(datos = datos,fctr = interaccion)
p2 <- fn.boxplot(datos = datos,fctr = interaccion)
pca.log <- prcomp(t(datos),center = T,scale. = T)
pca.var <- 100*(pca.log$sdev^2/sum(pca.log$sdev^2))
p3<-graphPCA.methods(pca.log$rotation[,1:2],
                     datos.common = datos.common,
                     varianzas=pca.var)+ggtitle("PCA")
pca.spearman <- princomp(cor(t(datos),method = "spearman"),fix_sign = T,cor=T)

pca.spearman.scores <- pca.spearman$scores
pca.spearman.var <- 100*(pca.spearman$sdev^2/sum(pca.spearman$sdev^2))
p4<-graphPCA.methods(pca.spearman.scores[,1:2],
                 datos.common = datos.common,
                 varianzas = pca.spearman.var)+ggtitle("PCA spearman")



plsda.log <- plsda(datos,Y=interaccion,scale = F)
plsda.var <-100*plsda.log$prop_expl_var$X
p5<-graphPCA.methods(datos=plsda.log$variates$X,
                     datos.common = datos.common,
                     varianzas = plsda.var)+ggtitle("PLS-DA")




p7<-ggarrange(p1,p2,common.legend = T)
annotate_figure(p7,bottom = "TMMwsp")
p8<-ggarrange(p3,p4,ncol = 2,common.legend = T)
annotate_figure(p8,bottom = "TMMwsp")
p9<-ggarrange(p5,p6,ncol = 2,common.legend = T)
p5
```

