---
title: Integration of omics and clinical data of the hormonal, metabolic, inflammatory,
  and oxidative response in the different macronutrients of the diet
author: "Edmond Geraud"
subtitle: '`r params$subtitulo`'
date: '`r format(Sys.Date(),"%e de %B, %Y")`' 
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    toc: true
  html_document:
    toc: true
    toc_depth: 2
header-includes:
  - \usepackage[english]{babel}
params:
  
  file_IP: Integromics_IPmarkers.xlsx
  file_metabolome: Integromics_Metabolome.xlsx
  file_general: Integromics_1.xlsx
  file_microbiota : integromics_microbiota.xlsx
  folder.data: ../../datos
  subtitulo: FINAL RESULTS
geometry: margin=2cm
---
```{r setup, include=FALSE}
require(knitr)
# include this code chunk as-is to set options
opts_chunk$set(comment = NA, prompt = TRUE, tidy = T, 
               fig.width = 7, fig.height = 7,echo = TRUE, 
               message = FALSE, warning = FALSE, cache=TRUE, fig_caption = TRUE)
Sys.setlocale("LC_TIME", "C")
```

## Load the libraries

```{r libraries}

list.of.packages <- c("xlsx","kableExtra","dplyr","ggplot2","egg","cowplot","patchwork","gridExtra","UsingR","car","lattice","ggpubr","ggbreak","GGally","reshape2","ggcorrplot","corrplot","rela","ggrepel","factoextra","chemometrics","sparsepca","vegan","ca","ade4","pls","OmicsPLS")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])] ## check for packages not installed 
if(length(new.packages)) install.packages(new.packages) ## install packages if necessary
res<-unlist(lapply(list.of.packages, require,character.only = T)) ## load packages needed for the session
# if(any(res==F)){
  list.of.packages[which(res==F)]  ## show those package if you have troubles to install.


list.of.bioc.packages <- c("phyloseq","pcaMethods","ropls","mixOmics","limma","DESeq2")
new.packages.bioc <- list.of.packages[!(list.of.bioc.packages %in% installed.packages()[,"Package"])]
if(length(new.packages.bioc)) BiocManager::install(new.packages.bioc)
res.bioc <- unlist(lapply(list.of.bioc.packages, require,character.only = T)) 
if(any(res.bioc==F)){
  list.of.packages[which(res.bioc==F)]  ## show those package if you have troubles to install.
}
```
# Funciones
```{r functions}


proces_bacteria <- function(datos,tipo){

  Order <- bacteria_phylum.abs$Order
  variables_in_bacteria <- general_data[general_data$Orden %in% Order , ]
  Sample<-variables_in_bacteria$Paciente


  if(tipo=="genera"){

    Order <- Order

    X <- datos[-1,-1]
    X<-apply(X,2,as.numeric)
    colnames(X) <- datos[1,2:ncol(datos)]
    rownames(X) <- Sample
    X.rel <- 100*X/rowSums(X)


  } else if(tipo=="phylum"){

    X <- datos[,-c(1:2)]
    colnames(X) <- colnames(datos)[3:ncol(datos)]
    rownames(X) <- Sample
    X.rel<-100*X/rowSums(X)
  }

  GROUP <-variables_in_bacteria$GROUP
  SEX <- variables_in_bacteria$SEX
  OBESE<- variables_in_bacteria$OBESE

  X<-as.data.frame(X)
  X.rel<-as.data.frame(X.rel)

  X$GROUP<-GROUP
  X$SEX<-SEX
  X$OBESE<-OBESE


  X.rel$GROUP<-GROUP
  X.rel$SEX<-SEX
  X.rel$OBESE<-OBESE

  return(list(abs=X,rel=X.rel))

}


low.count.removal = function(
                        data, # OTU count data frame of size n (sample) x p (OTU)
                        percent=0.01 # cutoff chosen
                        ){
    keep.otu = which(colSums(data)*100/(sum(colSums(data))) > percent)
    data.filter = data[,keep.otu]
    return(list(data.filter = data.filter, keep.otu = keep.otu))
}

low.count.removal.counts = function(
                        data, # OTU count data frame of size n (sample) x p (OTU)
                        abundance=10 # cutoff chosen
                        ){
    keep.otu <- which(colSums(data)>abundance)
    data.filter <-data[,keep.otu]
    return(list(data.filter = data.filter, keep.otu = keep.otu))
}

relative.fn <-function(data){
  
  return(100*data/rowSums(data))
  
}


common.bacteria <- function(datos_i,datos.comunes){
  

  ## retorna los indices con las variables comunes

  interaccion <- interaction(datos.comunes$GROUP,datos.comunes$OBESE)
  datos.comunes$interaccion <- interaccion
  grupo <-levels(interaccion)
  w<-vector(mode="list",length=6)
  names(w)<-levels(interaccion)
  for(i in 1:6){
   w[[names(w)[i]]]<-which(apply(as.matrix(datos_i[
    rownames(datos_i)%in%
      datos.comunes[datos.comunes$interaccion==grupo[i],"Paciente"],]),2,function(x) sum(x)==0))
  }

  w2<-vector(mode="list",length=6)
  names(w2)<-levels(interaccion)
  for(i in 1:6){
   w2[[names(w2)[i]]]<-which(apply(as.matrix(datos_i[
    rownames(datos_i)%in%
      datos.comunes[datos.comunes$interaccion==grupo[i],"Paciente"],]),2,function(x) sum(x)>0))
    }

  w3<-lapply(w2, names)
  common_variables <-Reduce(intersect,x = w3)

  return(common_variables)
}

fn.densidad <- function(datos,fctr){
  column_var=c("interaccion","variable","value")
  columnas <- colnames(datos)
  filas <- rownames(datos)
  X<-data.frame(apply(datos, 2, as.numeric))
  colnames(X)<-columnas
  rownames(X)<-filas
  X$fctr<-fctr
  X.melt <- melt(X)
  colnames(X.melt)<-column_var
  p<-ggplot(data=X.melt,aes(x=value,color=interaccion))+geom_density()
  
  return(p)
}

fn.boxplot<-function(datos,fctr){
  column_var=c("interaccion","variable","value")
  columnas <- colnames(datos)
  filas <- rownames(datos)
  X<-data.frame(apply(datos, 2, as.numeric))
  colnames(X)<-columnas
  rownames(X)<-filas
  X$fctr<-fctr
  X.melt <- melt(X)
  colnames(X.melt)<-column_var
  
  p<-ggplot(X.melt,aes(x=interaccion,y=value,fill=interaccion))+geom_boxplot()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  return(p)
}


graphPCA.methods <- function(datos,datos.common,varianzas){
  
  
  sujetos <-as.data.frame(datos)
  sexo<-datos.common$GROUP
  obesidad<-datos.common$OBESE
  colnames(sujetos)<-c("PC1","PC2")
  p1<-ggplot(sujetos,aes(x=PC1, y=PC2)) +
     geom_hline(yintercept = 0, color = "gray70") +
     geom_vline(xintercept = 0, color = "gray70") +
     geom_point(aes(color = sexo,shape=obesidad,fill=obesidad), alpha = 0.55, size = 3) +
     coord_cartesian(xlim = c(min(sujetos[,1]),max(sujetos[,1]))) +
     scale_color_manual(values=c("black","red","green"))+scale_shape_manual(values=c(17,15))+xlab(paste("PC1",round(varianzas[1],2),"%"))+ylab(paste("PC2",round(varianzas[2],2),"%"))
  return(p1)
}
graphPCA.methods2 <- function(datos,datos.common,varianzas){
  
  
  sujetos <-as.data.frame(datos)
  sexo<-datos.common$GROUP
  obesidad<-datos.common$OBESE
  colnames(sujetos)<-c("PC1","PC2")
  p1<-ggplot(sujetos,aes(x=PC1, y=PC2)) +
     geom_hline(yintercept = 0, color = "gray70") +
     geom_vline(xintercept = 0, color = "gray70") +
     geom_point(aes(color = obesidad,shape=sexo,fill=obesidad), alpha = 0.55, size = 3) +
     coord_cartesian(xlim = c(min(sujetos[,1]),max(sujetos[,1]))) +
     scale_color_manual(values=c("black","red"))+scale_shape_manual(values=c(17,15,14))+xlab(paste("PC1",round(varianzas[1],2),"%"))+ylab(paste("PC2",round(varianzas[2],2),"%"))
  return(p1)
}

```


# Cargamos los datos

## Datos clinicos

```{r}
general_data <- read.xlsx(file.path(params$folder.data,params$file_general),sheetIndex =1)
general_data$SEX <- factor(general_data$SEX, levels = c(0, 2), labels = c("Female", "Male"))
general_data$OBESE <- factor(general_data$OBESE, levels = c(0, 1), labels = c("No Obese", "Obese"))
general_data$GROUP <- factor(general_data$GROUP, levels = c(0, 1, 2), labels = c("Female",  "PCOS", "Male"))


```


## Datos IP

```{r}

IP_file<-file.path(params$folder.data,file=params$file_IP) ## file path of the data
IP.tmp<-read.xlsx(IP_file,sheetIndex = 1)
IP<-IP.tmp[,-7] # remove the missin values column
IP$SEX <- factor(IP$SEX,levels=c(0,2),labels = c("Female","Male"))
IP$OBESE <- factor(IP$OBESE,levels=c(0,1),labels = c("No Obese","Obese"))
IP$GROUP <- factor(IP$GROUP,levels=c(0,1,2),labels = c("Female","PCOS","Male"))

```

## Datos Metaboloma

```{r}
metabolome_file<-file.path(params$folder.data,file=params$file_metabolome) ## file path of the data
metabolome.tmp<-read.xlsx(metabolome_file,sheetIndex = 1) # read the data
metabolites_name <- read.xlsx(metabolome_file,sheetIndex = 2)$Metabolitos # read the names of the metabolites
any(is.na(metabolome.tmp)) # there is a row of missing values.
metabolome<-metabolome.tmp[-nrow(metabolome.tmp),] ## remove the missing value row
metabolome$GROUP<-general_data$GROUP ## add GROUP variable 
metabolome$OBESE <- general_data$OBESE ## add OBESE variables.

```


## Datos microbioma

```{r}
bacteria_phylum.abs <- as.data.frame(read.xlsx(file.path(params$folder.data,params$file_microbiota), sheetIndex = 1))
bacteria_genera.abs <- as.data.frame(t(read.xlsx(file.path(params$folder.data,params$file_microbiota), sheetIndex = 3)))
phylum.list <- proces_bacteria(bacteria_phylum.abs,tipo="phylum")
genera.list <- proces_bacteria(bacteria_genera.abs,tipo="genera")

phylum.abs <- phylum.list$abs;phylum.rel<-phylum.list$rel
genera.abs <- genera.list$abs;genera.rel<-genera.list$rel

```

# Preparando los datos para la integracion
```{r}
variables_in_bacteria <- general_data[general_data$Orden %in% bacteria_phylum.abs$Order, ]

```

## Datos Clinicos
**NOTE we are going to include also on both levels of integration body measures**
We exclude ISI because it is measured after the load of glucose, the mean values, to exclude redundancy of the data, ant the postprandial levels. 


```{r}
sujetos <- variables_in_bacteria$Paciente
Clin.tmp <- variables_in_bacteria[,-c(1,2,3,4,5)] # get rid off orden, paciente, sex, group and obese
auc_clin <- colnames(Clin.tmp)[grep("AUC",colnames(Clin.tmp))]
mean_basal_clin<-colnames(Clin.tmp)[grep("_B$",colnames(Clin.tmp))]
sog <- colnames(Clin.tmp)[grep("^SO[GLP]",colnames(Clin.tmp))]
variables_clin <- colnames(Clin.tmp)
testosterona <-colnames(Clin.tmp)[grep("TEST",colnames(Clin.tmp))]
interr<-colnames(Clin.tmp)[grep("interaccion",colnames(Clin.tmp))]
ratio <- colnames(Clin.tmp)[grep("Ratio",colnames(Clin.tmp))]
w<-c(which(variables_clin %in% auc_clin),which(variables_clin %in% mean_basal_clin),which(variables_clin %in% sog),which(variables_clin %in% testosterona),
which(variables_clin %in% interr),which(variables_clin %in% ratio))
Clin <- Clin.tmp[,-w] 
rownames(Clin) <- sujetos
colnames(Clin)
```


## Datos IP
```{r}
IP.tmp <- IP[IP$Paciente %in% variables_in_bacteria$Paciente,]
redundant_variables <- c("Orden","Paciente","SEX","GROUP","OBESE","BMI")
variables_ip <- colnames(IP.tmp)
auc_ip <- variables_ip[grep("AUC",variables_ip)]
mean_ip <- variables_ip[grep("_0$",variables_ip)]
w <- c(which(variables_ip %in% redundant_variables),which(variables_ip %in% auc_ip),which(variables_ip %in% mean_ip))
IP.basal <- IP.tmp[,-w]
rownames(IP.basal) <- sujetos

```

## Metabolome data (37)

```{r}
metabolome.basal.tmp <- metabolome[metabolome$Order %in% variables_in_bacteria$Orden,]
redundant_data_meta <- c("Order","Sample","GROUP","OBESE")
variables_meta <- colnames(metabolome.basal.tmp)
mean_meta <- variables_meta[grep("_[GLP]0$",variables_meta)]
auc_meta <- variables_meta[grep("AU[CG]",variables_meta)]
w <- c(which(variables_meta %in% redundant_data_meta),which(variables_meta %in% mean_meta),which(variables_meta %in% auc_meta))
metabolome.basal_mean <- metabolome.basal.tmp[,-w]
rownames(metabolome.basal_mean) <- sujetos
colnames(metabolome.basal_mean)
```
## Datos microbioma

### Datos genero

```{r}

genera.abs.tmp <- genera.abs
redundant_data_ph <- c("GROUP","SEX","OBESE")
genus.tmp.tmp <- genera.abs.tmp[,-c(ncol(genera.abs.tmp)-2,ncol(genera.abs.tmp)-1,ncol(genera.abs.tmp))]
genus <- genus.tmp.tmp[,colSums(genus.tmp.tmp)>0]
rownames(genus) <- sujetos
## removal of genus unidentified

genus<-genus[,-grep("^ud-",colnames(genus))]

common_genus <-common.bacteria(genus,variables_in_bacteria)

genus<-genus[,common_genus]
genus.abs <- low.count.removal.counts(genus,10)$data.filter

```

# Datos que necesitamos

```{r}

pacientes <- variables_in_bacteria$Paciente
grupo<-variables_in_bacteria$GROUP
obesidad <- factor(gsub(" ",".",variables_in_bacteria$OBESE))
interaccion<-factor(paste(obesidad,grupo,sep="_"))
datos.common <- variables_in_bacteria[,c("GROUP","OBESE","Paciente")]
columnData <- data.frame(interaccion=interaccion)
rownames(columnData) <- pacientes

```

# Descriptiva

## Metaboloma

```{r}
met37<-t(metabolome.basal_mean)
met37_norm <- met37/rowSums(met37)
met37_log <- t(apply(met37_norm,1,log))
X<-(voom(met37_norm,normalize.method = "quantile")$E)
p1<-fn.densidad((t(X)),interaccion)
p2<-fn.boxplot(t(X),interaccion)
pcx <- prcomp(t(X),scale=F,center=T)
varianzas <- 100*(pcx$sdev^2/sum(pcx$sdev^2))
p3<-graphPCA.methods(pcx$x,datos.common,varianzas)+ggtitle("PCA")

U3 <- pcx$x
apply(U3, 2, function(x) which( abs(x - mean(x)) > (6 * sd(x)) ))



pcx.cor <- princomp(x=cor((X),method = "spearman"),cor=T)
varianzas <- 100*(pcx.cor$sdev^2/sum(pcx.cor$sdev^2))
p4<-graphPCA.methods(pcx.cor$scores[,1:2],datos.common,varianzas)+ggtitle("PCA spearman")

plsda.met <- plsda(t(X),interaccion,scale = T)
p5<-graphPCA.methods(plsda.met$variates$X,datos.common,varianzas)+ggtitle("PLS-DA")

p7 <-graphPCA.methods2(pcx$x,datos.common,varianzas)+ggtitle("PCA")

euc <- vegdist(t(X),method = "euclidean")
points<-cmdscale(euc)
p6<-graphPCA.methods(points,datos.common,varianzas)+ggtitle("MDS")
ggarrange(p1,p2,p3,p7,
          common.legend = T)
ggarrange(p3,p4,p5,p6,ncol=2,nrow=2,common.legend = T)
ggarrange(p3,p7)
(res<-adonis2(euc~obesidad*grupo))

```




```{r}
fviz_pca_var(pcx)
```

```{r}
fviz_eig(pcx, addlabels = TRUE, ylim = c(0, 50))

```
```{r}
var.pca <- get_pca_var(pcx)
corrplot::corrplot(var.pca$cos2, is.corr=FALSE)
```
```{r}
fviz_contrib(pcx, choice = "var", axes = 1, top = 10)
```
```{r}
fviz_contrib(pcx, choice = "var", axes = 2, top = 10)
```

## Microbioma

Segun la guia del paper que me dio Susana, tenemos 2 opciones

```{r}
genera.abs.tmp <- genera.abs
redundant_data_ph <- c("GROUP","SEX","OBESE")
genus.tmp.tmp <- genera.abs.tmp[,-c(ncol(genera.abs.tmp)-2,ncol(genera.abs.tmp)-1,ncol(genera.abs.tmp))]
genus <- genus.tmp.tmp[,colSums(genus.tmp.tmp)>0]
rownames(genus) <- sujetos
## removal of genus unidentified

genus<-genus[,-grep("^ud-",colnames(genus))]

common_genus <-common.bacteria(genus,variables_in_bacteria)

genus<-genus[,common_genus]
genus.abs <- low.count.removal.counts(genus,10)$data.filter
library(zCompositions)
# genus.imput <-cmultRepl(X = (genus.abs),label = 0,output = "p-counts")
# genus.G <-(logratio.transfo(t(genus.imput),logratio = "CLR"))

genus.imput <-cmultRepl(X = (genus.abs),label = 0,output = "p-counts")

G <-matrix(t(logratio.transfo(t(genus.imput),logratio = "CLR")),ncol=ncol(genus.abs),nrow(genus.abs))

X<-t(G)
rownames(X) <- colnames(genus.abs)
colnames(X) <- rownames(genus.abs)
p1<-fn.densidad(t((X)),interaccion)
p2<-fn.boxplot(t(X),interaccion)
pcx <- prcomp(t(X),scale=F,center=T)

varianzas <- 100*(pcx$sdev^2/sum(pcx$sdev^2))
p3<-graphPCA.methods(pcx$x,datos.common,varianzas)+ggtitle("PCA")

U3 <- pcx$x
apply(U3, 2, function(x) which( abs(x - mean(x)) > (6 * sd(x)) ))



pcx.cor <- princomp(x=cor((X),method = "spearman"),cor=T)
varianzas <- 100*(pcx.cor$sdev^2/sum(pcx.cor$sdev^2))
p4<-graphPCA.methods(pcx.cor$scores[,1:2],datos.common,varianzas)+ggtitle("PCA spearman")

plsda.met <- plsda(t(X),interaccion,scale = T)
p5<-graphPCA.methods(plsda.met$variates$X,datos.common,varianzas)+ggtitle("PLS-DA")


euc <- vegdist(t(X),method = "euclidean")
points<-cmdscale(euc)
p6<-graphPCA.methods(points,datos.common,varianzas)+ggtitle("MDS")
ggarrange(p1,p2,ncol=2,common.legend = T)
ggarrange(p3,p4,p5,p6,ncol=2,nrow=2,common.legend = T)

(res<-adonis2(t(X)~obesidad*grupo,method = "euclidean"))
(res<-adonis2(plsda.met$X~obesidad*grupo,method = "euclidean"))
```



```{r}
varianzas <- 100*(pcx$sdev^2/sum(pcx$sdev^2))

p3<-graphPCA.methods(pcx$x,datos.common,varianzas)+ggtitle("PCA")
p4<-graphPCA.methods2(pcx$x,datos.common,varianzas)+ggtitle("PCA")
ggarrange(p3,p4,common.legend = T)
```



```{r}
fviz_eig(pcx, addlabels = TRUE, ylim = c(0, 50))

```
```{r}
var.pca <- get_pca_var(pcx)
corrplot::corrplot(var.pca$cos2, is.corr=FALSE)
```
```{r}
fviz_contrib(pcx, choice = "var", axes = 1, top = 10)
```
```{r}
fviz_contrib(pcx, choice = "var", axes = 2, top = 10)
```
```{r}
met37_log <- t(apply(met37_norm,1,log))
metaboloma<-(voom(met37_norm,normalize.method = "quantile")$E)
microbioma <- (G)
colnames(microbioma) <- colnames(genus.abs)
rownames(microbioma) <- rownames(genus.abs)
microbioma <- t(microbioma)
```

```{r}
dim(metaboloma)
dim(microbioma)
```



```{r}
G<-t(G)
```

## Integracion
```{r}
library(limma)
library(MOFA2)
metaboloma<-(voom(met37_norm,normalize.method = "quantile")$E)

colnames(G) <- pacientes
rownames(G) <- colnames(genus.abs)
datos <- list(metaboloma=(metaboloma),microbioma=(G))
grupos <- data.frame(obesidad=obesidad,grupo=grupo)
rownames(grupos) <- pacientes
MOFAobject <-create_mofa_from_matrix(datos,groups =interaccion)
samples_metadata(MOFAobject)<- data.frame(sample=pacientes,obesidad=(obesidad),grupo=(grupo),group=interaccion)
rownames(samples_metadata(MOFAobject))<-pacientes
plot_data_overview(MOFAobject)


```


```{r}
data_opts <- get_default_data_options(MOFAobject)
data_opts$scale_views <-T
data_opts$scale_groups <- T
# data_opts$groups <- data.frame(grupo=grupo,obesidad=obesidad)
head(data_opts)

```


```{r}
model_opts <- get_default_model_options(MOFAobject)
model_opts$likelihoods=c("gaussian","gaussian")
model_opts$num_factors=15
head(model_opts)
model_opts$spikeslab_factors<-F
model_opts$ard_factors <-T
```
```{r}
train_opts <- get_default_training_options(MOFAobject)
head(train_opts)

```


```{r}
# outfile = file.path(getwd(),paste0(seq(1:10),"_model.hdf5"))
  MOFAobject <- prepare_mofa(
  object = MOFAobject,
  data_options = data_opts,
  model_options = model_opts,
  training_options = train_opts
)
outfile <- file.path(getwd(),"model.hdf5")
reticulate::use_python("/usr/bin/python3")
list.trained <- vector("list",length = 10)


MOFAobject.trained <- run_mofa(MOFAobject, outfile,use_basilisk = F)
MOFAobject.trained@covariates<-data.frame(obesidad=as.numeric(obesidad),grupo=as.numeric(grupo))
MOFAobject.trained
```

```{r}
plot_variance_explained(MOFAobject.trained, x="group", y="factor")
```


```{r}
plot_variance_explained(
  MOFAobject.trained, 
  x = "group", 
  y = "factor", 
  plot_total = T
)[[2]]
```

```{r}
plot_variance_explained_per_feature(
  MOFAobject.trained, 
  view = "microbioma",
)
```

