---
title: Integration of omics and clinical data of the hormonal, metabolic, inflammatory,
  and oxidative response in the different macronutrients of the diet
author: "Edmond Geraud"
subtitle: '`r params$subtitulo`'
date: '`r format(Sys.Date(),"%e de %B, %Y")`' 
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    toc: true
  html_document:
    toc: true
    toc_depth: 2
header-includes:
  - \usepackage[english]{babel}
params:
  
  file_IP: Integromics_IPmarkers.xlsx
  file_metabolome: Integromics_Metabolome.xlsx
  file_general: Integromics_1.xlsx
  file_microbiota : integromics_microbiota.xlsx
  folder.data: ../../../datos
  subtitulo: FINAL RESULTS
geometry: margin=2cm
---
```{r setup, include=FALSE}
require(knitr)
# include this code chunk as-is to set options
opts_chunk$set(comment = NA, prompt = TRUE, tidy = T, 
               fig.width = 7, fig.height = 7,echo = TRUE, 
               message = FALSE, warning = TRUE, cache=TRUE, fig_caption = TRUE)
Sys.setlocale("LC_TIME", "C")
```

## Load the libraries

```{r libraries}

list.of.packages <- c("xlsx","kableExtra","dplyr","ggplot2","egg","cowplot","patchwork","gridExtra","UsingR","car","lattice","ggpubr","ggbreak","GGally","reshape2","ggcorrplot","corrplot","rela","ggrepel","factoextra","chemometrics","sparsepca","vegan","ca","ade4","pls","OmicsPLS","psych")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])] ## check for packages not installed 
if(length(new.packages)) install.packages(new.packages) ## install packages if necessary
res<-unlist(lapply(list.of.packages, require,character.only = T)) ## load packages needed for the session
# if(any(res==F)){
  list.of.packages[which(res==F)]  ## show those package if you have troubles to install.


list.of.bioc.packages <- c("phyloseq","pcaMethods","ropls","mixOmics","limma","DESeq2")
new.packages.bioc <- list.of.packages[!(list.of.bioc.packages %in% installed.packages()[,"Package"])]
if(length(new.packages.bioc)) BiocManager::install(new.packages.bioc)
res.bioc <- unlist(lapply(list.of.bioc.packages, require,character.only = T)) 
if(any(res.bioc==F)){
  list.of.packages[which(res.bioc==F)]  ## show those package if you have troubles to install.
}
```
# Funciones
```{r functions}


proces_bacteria <- function(datos,tipo){

  Order <- bacteria_phylum.abs$Order
  variables_in_bacteria <- general_data[general_data$Orden %in% Order , ]
  Sample<-variables_in_bacteria$Paciente


  if(tipo=="genera"){

    Order <- Order

    X <- datos[-1,-1]
    X<-apply(X,2,as.numeric)
    colnames(X) <- datos[1,2:ncol(datos)]
    rownames(X) <- Sample
    X.rel <- 100*X/rowSums(X)


  } else if(tipo=="phylum"){

    X <- datos[,-c(1:2)]
    colnames(X) <- colnames(datos)[3:ncol(datos)]
    rownames(X) <- Sample
    X.rel<-100*X/rowSums(X)
  }

  GROUP <-variables_in_bacteria$GROUP
  SEX <- variables_in_bacteria$SEX
  OBESE<- variables_in_bacteria$OBESE

  X<-as.data.frame(X)
  X.rel<-as.data.frame(X.rel)

  X$GROUP<-GROUP
  X$SEX<-SEX
  X$OBESE<-OBESE


  X.rel$GROUP<-GROUP
  X.rel$SEX<-SEX
  X.rel$OBESE<-OBESE

  return(list(abs=X,rel=X.rel))

}


low.count.removal = function(
                        data, # OTU count data frame of size n (sample) x p (OTU)
                        percent=0.01 # cutoff chosen
                        ){
    keep.otu = which(colSums(data)*100/(sum(colSums(data))) > percent)
    data.filter = data[,keep.otu]
    return(list(data.filter = data.filter, keep.otu = keep.otu))
}

low.count.removal.counts = function(
                        data, # OTU count data frame of size n (sample) x p (OTU)
                        abundance=10 # cutoff chosen
                        ){
    keep.otu <- which(colSums(data)>abundance)
    data.filter <-data[,keep.otu]
    return(list(data.filter = data.filter, keep.otu = keep.otu))
}

relative.fn <-function(data){
  
  return(100*data/rowSums(data))
  
}


common.bacteria <- function(datos_i,datos.comunes){
  

  ## retorna los indices con las variables comunes

  interaccion <- interaction(datos.comunes$GROUP,datos.comunes$OBESE)
  datos.comunes$interaccion <- interaccion
  grupo <-levels(interaccion)
  w<-vector(mode="list",length=6)
  names(w)<-levels(interaccion)
  for(i in 1:6){
   w[[names(w)[i]]]<-which(apply(as.matrix(datos_i[
    rownames(datos_i)%in%
      datos.comunes[datos.comunes$interaccion==grupo[i],"Paciente"],]),2,function(x) sum(x)==0))
  }

  w2<-vector(mode="list",length=6)
  names(w2)<-levels(interaccion)
  for(i in 1:6){
   w2[[names(w2)[i]]]<-which(apply(as.matrix(datos_i[
    rownames(datos_i)%in%
      datos.comunes[datos.comunes$interaccion==grupo[i],"Paciente"],]),2,function(x) sum(x)>0))
    }

  w3<-lapply(w2, names)
  common_variables <-Reduce(intersect,x = w3)

  return(common_variables)
}

fn.densidad <- function(datos,fctr){
  column_var=c("interaccion","variable","value")
  columnas <- colnames(datos)
  filas <- rownames(datos)
  X<-data.frame(apply(datos, 2, as.numeric))
  colnames(X)<-columnas
  rownames(X)<-filas
  X$fctr<-fctr
  X.melt <- melt(X)
  colnames(X.melt)<-column_var
  p<-ggplot(data=X.melt,aes(x=value,color=interaccion))+geom_density()
  
  return(p)
}

fn.boxplot<-function(datos,fctr){
  column_var=c("interaccion","variable","value")
  columnas <- colnames(datos)
  filas <- rownames(datos)
  X<-data.frame(apply(datos, 2, as.numeric))
  colnames(X)<-columnas
  rownames(X)<-filas
  X$fctr<-fctr
  X.melt <- melt(X)
  colnames(X.melt)<-column_var
  
  p<-ggplot(X.melt,aes(x=interaccion,y=value,fill=interaccion))+geom_boxplot()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  return(p)
}


graphPCA.methods <- function(datos,datos.common,varianzas){
  
  
  sujetos <-as.data.frame(datos)
  sexo<-datos.common$GROUP
  obesidad<-datos.common$OBESE
  colnames(sujetos)<-c("PC1","PC2")
  p1<-ggplot(sujetos,aes(x=PC1, y=PC2)) +
     geom_hline(yintercept = 0, color = "gray70") +
     geom_vline(xintercept = 0, color = "gray70") +
     geom_point(aes(color = sexo,shape=obesidad,fill=obesidad), alpha = 0.55, size = 3) +
     coord_cartesian(xlim = c(min(sujetos[,1]),max(sujetos[,1]))) +
     scale_color_manual(values=c("black","red","green"))+scale_shape_manual(values=c(17,15))+xlab(paste("PC1",round(varianzas[1],2),"%"))+ylab(paste("PC2",round(varianzas[2],2),"%"))
  return(p1)
}
graphPCA.methods2 <- function(datos,datos.common,varianzas){
  
  
  sujetos <-as.data.frame(datos)
  sexo<-datos.common$GROUP
  obesidad<-datos.common$OBESE
  colnames(sujetos)<-c("PC1","PC2")
  p1<-ggplot(sujetos,aes(x=PC1, y=PC2)) +
     geom_hline(yintercept = 0, color = "gray70") +
     geom_vline(xintercept = 0, color = "gray70") +
     geom_point(aes(color = obesidad,shape=sexo,fill=obesidad), alpha = 0.55, size = 3) +
     coord_cartesian(xlim = c(min(sujetos[,1]),max(sujetos[,1]))) +
     scale_color_manual(values=c("black","red"))+scale_shape_manual(values=c(17,15,14))+xlab(paste("PC1",round(varianzas[1],2),"%"))+ylab(paste("PC2",round(varianzas[2],2),"%"))
  return(p1)
}
plot.methods2 <- function(factores,datos.common){
  
  
  sujetos <-data.frame(Factor1 = factores[,1], Factor2=factores[,2])
  sexo<-datos.common$GROUP
  obesidad<-datos.common$OBESE
  p1<-ggplot(sujetos,aes(x=Factor1, y=Factor2)) +
     geom_hline(yintercept = 0, color = "gray70") +
     geom_vline(xintercept = 0, color = "gray70") +
     geom_point(aes(color = sexo,shape=obesidad,fill=obesidad), alpha = 0.55, size = 3)  +
     scale_color_manual(values=c("black","red","green"))+scale_shape_manual(values=c(17,15))
  return(p1)
}

plot.methods <- function(factores,datos.common){
  
  
  sujetos <-data.frame(Sujetos=1:length(factores),Factor=factores)
  sexo<-datos.common$GROUP
  p1<-ggplot(sujetos,aes(x=Sujetos, y=Factor)) +
     geom_hline(yintercept = 0, color = "gray70") +
     geom_vline(xintercept = 0, color = "gray70") +
     geom_point(aes(color = sexo,fill=grupo), alpha = 0.55, size = 3)  +
     scale_color_manual(values=c("black","red","green"))+scale_shape_manual(values=c(17,15))
  return(p1)
}


plot.methods2.ob <- function(factores,datos.common){
  
  
  sujetos <-data.frame(Factor1 = factores[,1], Factor2=factores[,2])
  sexo<-datos.common$GROUP
  obesidad<-datos.common$OBESE
  p1<-ggplot(sujetos,aes(x=Factor1, y=Factor2)) +
     geom_hline(yintercept = 0, color = "gray70") +
     geom_vline(xintercept = 0, color = "gray70") +
     geom_point(aes(color = obesidad,shape=grupo,fill=grupo), alpha = 0.55, size = 3)  +
     scale_color_manual(values=c("black","red"))+scale_shape_manual(values=c(17,15,13))
  return(p1)
}

extract_top_weights <- function(modelo,view,factors,thresh=0.5){
  
  p1<-plot_top_weights(modelo,factors=factors,view=view,scale=T)
top_weights <-as.character(unique(p1$data$feature[p1$data$value>=thresh]))
return(top_weights)
}
fun_restore <- function(modelo,grupo){
  factors <- get_factors(modelo, as.data.frame = T,scale = F,groups = grupo)

weights <- rbind(get_weights(modelo,as.data.frame=T,scale=F,views="metaboloma"),
                 get_weights(modelo,as.data.frame=T,scale=F,views="microbioma"))
                 
factores.recast <-as.data.frame(reshape2::recast(data=factors,sample~factor))
rownames(factores.recast) <- factores.recast$sample
factores.recast <- factores.recast[,-1]
weights.recast <-as.data.frame(reshape2::recast(data=weights,feature~factor))
rownames(weights.recast) <- weights.recast$feature
weights.recast <- weights.recast[,-1]
weights.recast$views <- c(rep("metabooloma",37),rep("microbioma",94))
Y <-(as.matrix(factores.recast) %*% t(weights.recast[,-ncol(weights.recast)]))
return(list(Y=Y,factores=factores.recast,pesos=weights.recast))
}
fn.explore <- function(datos.in){
  datos <- as.data.frame(datos.in)
grupo<-variables_in_bacteria$GROUP
obesidad <- factor(gsub(" ",".",variables_in_bacteria$OBESE))
interaccion<-factor(paste(obesidad,grupo,sep="_"))
datos$grupo <- grupo
datos$obesidad <- obesidad
datos$interaccion <- interaccion
datos$sexo <-variables_in_bacteria$SEX
descarte <--c(dim(datos.in)[2]:(dim(datos.in)[2]+4))
#### gruposde interes
femalesOnly <- datos[datos$sexo=="Female",descarte]
malesOnly <- datos[datos$sexo=="Male",descarte]
females <- datos[datos$grupo=="Female",descarte]
males <- datos[datos$grupo=="Male",descarte]
pcos <- datos[datos$grupo=="PCOS",descarte]
obesos <- datos[datos$obesidad=="Obese",descarte]
No.Obesos <- datos[datos$obesidad=="No.Obese",descarte]
#### interacciones
females.no.obese <- datos[datos$interaccion=="No.Obese_Female",descarte]
females.obese <- datos[datos$interaccion=="Obese_Female",descarte]
males.no.obese <-  datos[datos$interaccion=="No.Obese_Male",descarte]
males.obese <-  datos[datos$interaccion=="Obese_Male",descarte]
pcos.no.obese <- datos[datos$interaccion=="No.Obese_PCOS",descarte]
pcos.obese <- datos[datos$interaccion=="Obese_PCOS",descarte]
out <- list(femalesOnly=femalesOnly,
            malesOnly=malesOnly,
            females=females,
            males=males,
            pcos=pcos,
            obesos=obesos,
            No.Obesos=No.Obesos,
            females.no.obese=females.no.obese,
            males.no.obese=males.no.obese,
            pcos.no.obese=pcos.no.obese,
            females.obese=females.obese,
            pcos.obese=pcos.obese,
           males.obese=males.obese,
           Y=datos[,descarte])
return(out)
}


fn.cor.res2 <- function(m1,m2,top_pesos.names,name1,name2){
## correlacion m1
m1.cor <-  cor(m1[,top_pesos.names],method = "spearman")
## cor test m1
m1.cor.p <-corr.test(m1.cor,method = "spearman")$p
# m1.cor.p[colnames(m1.cor.p) %in% colnames(metaboloma),colnames(m1.cor.p) %in% colnames(metaboloma)]<-1
# m1.cor.p[colnames(m1.cor.p) %in% colnames(genus.abs),colnames(m1.cor.p) %in% colnames(genus.abs)]<-1
m1.cor.p[lower.tri(m1.cor.p)]<-1
diag(m1.cor.p)<-1

m2.cor <-  cor(m2[,top_pesos.names],method = "spearman")
m2.cor.p <-corr.test(m2.cor,method = "spearman")$p
# m2.cor.p[colnames(m2.cor.p) %in% colnames(metaboloma),colnames(m2.cor.p) %in% colnames(metaboloma)]<-1
# m2.cor.p[colnames(m2.cor.p) %in% colnames(genus.abs),colnames(m2.cor.p) %in% colnames(genus.abs)]<-1
m2.cor.p[lower.tri(m2.cor.p)]<-1
diag(m2.cor.p)<-1



corrplot::corrplot(m1.cor,p.mat = m1.cor.p,sig.level = 0.05,insig = "label_sig",title = name1,mar = c(0,0,1,0))
corrplot::corrplot(m2.cor,p.mat = m2.cor.p,sig.level = 0.05,insig = "label_sig",title = name2,mar = c(0,0,1,0))
females.only.cor <- cor(m1[,top_pesos.names])
males.only.cor <- cor(m2[,top_pesos.names])
pls.obj<-pls(t(m1[,top_pesos.names]),t(m2[,top_pesos.names]),ncomp = 1)
cortest <-cor.test(pls.obj$variates$X,pls.obj$variates$Y,method = "spearman")$p.value
corval<-cor(pls.obj$variates$X,pls.obj$variates$Y,method = "spearman")

manteltest <-mantel.test(females.only.cor,males.only.cor,nperm = 10000)$p
wilcoxtest <- wilcox.test(pls.obj$X,pls.obj$Y)$p.value

print(paste0(sum(m1.cor.p<0.05),"m1"))
print(paste0(sum(m2.cor.p<0.05),"m1"))

return(c(cortest=cortest,manteltest=manteltest,wilcoxtest=wilcoxtest,corval = corval))
}

fn.cor.res1 <- function(m1,top_pesos.names,name1){
## correlacion m1
m1.cor <-  cor(m1[,top_pesos.names],method = "spearman")
## cor test m1
m1.cor.p <-corr.test(m1.cor,method = "spearman")$p
# m1.cor.p[colnames(m1.cor.p) %in% colnames(metaboloma),colnames(m1.cor.p) %in% colnames(metaboloma)]<-1
# m1.cor.p[colnames(m1.cor.p) %in% colnames(genus.abs),colnames(m1.cor.p) %in% colnames(genus.abs)]<-1
m1.cor.p[lower.tri(m1.cor.p)]<-1
diag(m1.cor.p)<-1

corrplot::corrplot(m1.cor,p.mat = m1.cor.p,sig.level = 0.05,insig = "label_sig",title = name1,mar = c(0,0,1,0))

}
fn.explore2 <- function(datos.in,variables_integracion){
  datos <- as.data.frame(datos.in)
interaccion<-factor(paste(variables_integracion$obesidad,variables_integracion$grupo,sep="_"))
datos$grupo <- variables_integracion$grupo
datos$obesidad <- variables_integracion$obesidad
datos$interaccion <- interaccion
datos$sexo <-variables_in_bacteria$SEX
descarte <--c(dim(datos.in)[2]:(dim(datos.in)[2]+4))
#### gruposde interes
femalesOnly <- datos[datos$sexo=="Female",descarte]
malesOnly <- datos[datos$sexo=="Male",descarte]
females <- datos[datos$grupo=="Female",descarte]
males <- datos[datos$grupo=="Male",descarte]
pcos <- datos[datos$grupo=="PCOS",descarte]
obesos <- datos[datos$obesidad=="Obese",descarte]
No.Obesos <- datos[datos$obesidad=="No.Obese",descarte]
#### interacciones
females.no.obese <- datos[datos$interaccion=="No.Obese_Female",descarte]
females.obese <- datos[datos$interaccion=="Obese_Female",descarte]
males.no.obese <-  datos[datos$interaccion=="No.Obese_Male",descarte]
males.obese <-  datos[datos$interaccion=="Obese_Male",descarte]
pcos.no.obese <- datos[datos$interaccion=="No.Obese_PCOS",descarte]
pcos.obese <- datos[datos$interaccion=="Obese_PCOS",descarte]
out <- list(femalesOnly=femalesOnly,
            malesOnly=malesOnly,
            females=females,
            males=males,
            pcos=pcos,
            obesos=obesos,
            No.Obesos=No.Obesos,
            females.no.obese=females.no.obese,
            males.no.obese=males.no.obese,
            pcos.no.obese=pcos.no.obese,
            females.obese=females.obese,
            pcos.obese=pcos.obese,
           males.obese=males.obese,
           Y=datos[,descarte])
return(out)
}
fun_restore <- function(modelo,grupo){
  factors <- get_factors(modelo, as.data.frame = T,scale = F,groups = grupo)

weights <- rbind(get_weights(modelo,as.data.frame=T,scale=F,views="metaboloma"),
                 get_weights(modelo,as.data.frame=T,scale=F,views="microbioma"))
                 
factores.recast <-as.data.frame(reshape2::recast(data=factors,sample~factor))
rownames(factores.recast) <- factores.recast$sample
factores.recast <- factores.recast[,-1]
weights.recast <-as.data.frame(reshape2::recast(data=weights,feature~factor))
rownames(weights.recast) <- weights.recast$feature
weights.recast <- weights.recast[,-1]
weights.recast$views <- c(rep("metabooloma",37),rep("microbioma",94))
Y <-(as.matrix(factores.recast) %*% t(weights.recast[,-ncol(weights.recast)]))
return(list(Y=Y,factores=factores.recast,pesos=weights.recast))
}

```


# PREPROCESADO

## Datos clinicos

```{r}
general_data <- read.xlsx(file.path(params$folder.data,params$file_general),sheetIndex =1)
general_data$SEX <- factor(general_data$SEX, levels = c(0, 2), labels = c("Female", "Male"))
general_data$OBESE <- factor(general_data$OBESE, levels = c(0, 1), labels = c("No Obese", "Obese"))
general_data$GROUP <- factor(general_data$GROUP, levels = c(0, 1, 2), labels = c("Female",  "PCOS", "Male"))


```


## Datos IP

```{r}

IP_file<-file.path(params$folder.data,file=params$file_IP) ## file path of the data
IP.tmp<-read.xlsx(IP_file,sheetIndex = 1)
IP<-IP.tmp[,-7] # remove the missin values column
IP$SEX <- factor(IP$SEX,levels=c(0,2),labels = c("Female","Male"))
IP$OBESE <- factor(IP$OBESE,levels=c(0,1),labels = c("No Obese","Obese"))
IP$GROUP <- factor(IP$GROUP,levels=c(0,1,2),labels = c("Female","PCOS","Male"))

```

## Datos Metaboloma

```{r}
metabolome_file<-file.path(params$folder.data,file=params$file_metabolome) ## file path of the data
metabolome.tmp<-read.xlsx(metabolome_file,sheetIndex = 1) # read the data
metabolites_name <- read.xlsx(metabolome_file,sheetIndex = 2)$Metabolitos # read the names of the metabolites
any(is.na(metabolome.tmp)) # there is a row of missing values.
metabolome<-metabolome.tmp[-nrow(metabolome.tmp),] ## remove the missing value row
metabolome$GROUP<-general_data$GROUP ## add GROUP variable 
metabolome$OBESE <- general_data$OBESE ## add OBESE variables.

```


## Datos microbioma

```{r}
bacteria_phylum.abs <- as.data.frame(read.xlsx(file.path(params$folder.data,params$file_microbiota), sheetIndex = 1))
bacteria_genera.abs <- as.data.frame(t(read.xlsx(file.path(params$folder.data,params$file_microbiota), sheetIndex = 3)))
phylum.list <- proces_bacteria(bacteria_phylum.abs,tipo="phylum")
genera.list <- proces_bacteria(bacteria_genera.abs,tipo="genera")

phylum.abs <- phylum.list$abs;phylum.rel<-phylum.list$rel
genera.abs <- genera.list$abs;genera.rel<-genera.list$rel

```

# Preparando los datos para la integracion
```{r}
variables_in_bacteria <- general_data[general_data$Orden %in% bacteria_phylum.abs$Order, ]

```

## Datos Clinicos
**NOTE we are going to include also on both levels of integration body measures**
We exclude ISI because it is measured after the load of glucose, the mean values, to exclude redundancy of the data, ant the postprandial levels. 


```{r}
sujetos <- variables_in_bacteria$Paciente
Clin.tmp <- variables_in_bacteria[,-c(1,2,3,4,5)] # get rid off orden, paciente, sex, group and obese
auc_clin <- colnames(Clin.tmp)[grep("AUC",colnames(Clin.tmp))]
mean_basal_clin<-colnames(Clin.tmp)[grep("_B$",colnames(Clin.tmp))]
sog <- colnames(Clin.tmp)[grep("^SO[GLP]",colnames(Clin.tmp))]
variables_clin <- colnames(Clin.tmp)
testosterona <-colnames(Clin.tmp)[grep("TEST",colnames(Clin.tmp))]
interr<-colnames(Clin.tmp)[grep("interaccion",colnames(Clin.tmp))]
ratio <- colnames(Clin.tmp)[grep("Ratio",colnames(Clin.tmp))]
w<-c(which(variables_clin %in% auc_clin),which(variables_clin %in% mean_basal_clin),which(variables_clin %in% sog),which(variables_clin %in% testosterona),
which(variables_clin %in% interr),which(variables_clin %in% ratio))
Clin <- Clin.tmp[,-w] 
rownames(Clin) <- sujetos
colnames(Clin)
```


## Datos IP
```{r}
IP.tmp <- IP[IP$Paciente %in% variables_in_bacteria$Paciente,]
redundant_variables <- c("Orden","Paciente","SEX","GROUP","OBESE","BMI")
variables_ip <- colnames(IP.tmp)
auc_ip <- variables_ip[grep("AUC",variables_ip)]
mean_ip <- variables_ip[grep("_0$",variables_ip)]
w <- c(which(variables_ip %in% redundant_variables),which(variables_ip %in% auc_ip),which(variables_ip %in% mean_ip))
IP.basal <- IP.tmp[,-w]
rownames(IP.basal) <- sujetos

```

## Metabolome data (37)

```{r}
metabolome.basal.tmp <- metabolome[metabolome$Order %in% variables_in_bacteria$Orden,]
redundant_data_meta <- c("Order","Sample","GROUP","OBESE")
variables_meta <- colnames(metabolome.basal.tmp)
mean_meta <- variables_meta[grep("_[GLP]0$",variables_meta)]
auc_meta <- variables_meta[grep("AU[CG]",variables_meta)]
w <- c(which(variables_meta %in% redundant_data_meta),which(variables_meta %in% mean_meta),which(variables_meta %in% auc_meta))
metabolome.basal_mean <- metabolome.basal.tmp[,-w]
rownames(metabolome.basal_mean) <- sujetos
colnames(metabolome.basal_mean)
```
## Datos microbioma

### Datos genero

```{r}

genera.abs.tmp <- genera.abs
redundant_data_ph <- c("GROUP","SEX","OBESE")
genus.tmp.tmp <- genera.abs.tmp[,-c(ncol(genera.abs.tmp)-2,ncol(genera.abs.tmp)-1,ncol(genera.abs.tmp))]
genus <- genus.tmp.tmp[,colSums(genus.tmp.tmp)>0]
rownames(genus) <- sujetos
## removal of genus unidentified

genus<-genus[,-grep("^ud-",colnames(genus))]

common_genus <-common.bacteria(genus,variables_in_bacteria)

genus<-genus[,common_genus]
genus.abs <- low.count.removal.counts(genus,10)$data.filter

```

# Datos que necesitamos

```{r}

pacientes <- variables_in_bacteria$Paciente
grupo<-variables_in_bacteria$GROUP
obesidad <- factor(gsub(" ",".",variables_in_bacteria$OBESE))
interaccion<-factor(paste(obesidad,grupo,sep="_"))
datos.common <- variables_in_bacteria[,c("GROUP","OBESE","Paciente")]
columnData <- data.frame(interaccion=interaccion)
rownames(columnData) <- pacientes

```


```{r}
met37<-t(metabolome.basal_mean)
met37_norm <- met37/rowSums(met37)
met37_log <- t(apply(met37_norm,1,log))
design <- model.matrix(~obesidad*grupo)
X<-(voom(met37_norm,normalize.method = "quantile",design = design)$E)

library(ALDEx2)
set.seed(123)
conds <- interaccion
x.all <- aldex.clr(t(genus.abs),conds)
```

# INTEGRACION

```{r}
library(MOFA2)
unique(strsplit2(list.files("../modelsMOFA_group/"),"_")[,3])

models <- lapply(list.files("../modelsMOFA_group/",full.names = T), load_model)

elbos <-compare_elbo(models,return_data = T)

model <-load_model(list.files("../modelsMOFA_group/",full.names = T)[which.max(elbos$ELBO)])
```

### Generalidades

```{r}
plot_factor_cor(model)

```
```{r}
ggarrange(plot_variance_explained(model, x="view", y="factor")+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)),
plot_variance_explained(model, y="factor", plot_total = T)[[2]]+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)))
calculate_variance_explained(model)
```

```{r}
model@covariates <-data.frame(obesidad=as.numeric(obesidad),grupo=as.numeric(grupo),edad=Clin$EDAD,BMI=Clin$BMI,WC=Clin$WC,WHR=Clin$WHR,SHGB=Clin$SHBG,hsCRP=Clin$hsCRP,TE2=Clin$Total_ESTR,E2=Clin$Free_ESTRA)

```


```{r}
correlate_factors_with_covariates(model, 
  covariates = c("grupo","edad","BMI","WC","WHR","SHGB","hsCRP","TE2","E2"), 
  plot="log_pval",groups = "all"
)
```


```{r}

p1 <- plot_factor(model,color_by  = "obesidad",add_boxplot = T,dodge=T,add_violin = T,rasterize = F,factors = 1,scale = T,groups = "PCOS")
p2 <- plot_factor(model,color_by  = "obesidad",group_by = "grupo",add_boxplot = T,dodge=T,add_violin = T,rasterize = F,factors = c(2),scale=T,groups = "Male")
p3 <- plot_factor(model,color_by  = "obesidad",group_by = "grupo",add_boxplot = T,dodge=T,add_violin = T,rasterize = F,factors = c(1),scale=T,groups = "Female")
ggarrange(p1,p2,p3,nrow=2,ncol=2)
```



### Correlacion covariables HOMBRES
```{r}
correlate_factors_with_covariates(model, 
  covariates = c("obesidad","edad","BMI","WC","WHR","SHGB","hsCRP","TE2","E2"), 
  plot="log_pval",groups = "Male")
```


```{r}
datagg <- as.data.frame(cbind(WHR=model@samples_metadata$WHR[model@samples_metadata$grupo=="Male"],Factor1=get_factors(model,factors=1,scale = T,as.data.frame = T,groups = "Male")$value))

  obesidad<-as.factor(model@samples_metadata$obesidad[model@samples_metadata$grupo=="Male"])
  p1<-ggplot(datagg,aes(x=WHR, y=Factor1)) +
     geom_hline(yintercept = 0, color = "gray70") +
     geom_vline(xintercept = 0, color = "gray70") +
     geom_point(aes(color = obesidad), alpha = 0.55, size = 3) +     scale_color_manual(values=c("black","red"))+ggtitle("Factor 1 Males WHR")+coord_cartesian(xlim = c(min(datagg[,1]),max(datagg[,1]))) 
  
 datagg <- as.data.frame(cbind(WC=model@samples_metadata$WC[model@samples_metadata$grupo=="Male"],Factor1=get_factors(model,factors=1,scale = T,as.data.frame = T,groups = "Male")$value))

  obesidad<-as.factor(model@samples_metadata$obesidad[model@samples_metadata$grupo=="Male"])
  p2<-ggplot(datagg,aes(x=WC, y=Factor1)) +
     geom_hline(yintercept = 0, color = "gray70") +
     geom_vline(xintercept = 0, color = "gray70") +
     geom_point(aes(color = obesidad), alpha = 0.55, size = 3) +     scale_color_manual(values=c("black","red"))+ggtitle("Factor 1 Males WC")+coord_cartesian(xlim = c(min(datagg[,1]),max(datagg[,1]))) 
  
 datagg <- as.data.frame(cbind(BMI=model@samples_metadata$BMI[model@samples_metadata$grupo=="Male"],Factor1=get_factors(model,factors=1,scale = T,as.data.frame = T,groups = "Male")$value))

  obesidad<-as.factor(model@samples_metadata$obesidad[model@samples_metadata$grupo=="Male"])
  p3<-ggplot(datagg,aes(x=BMI, y=Factor1)) +
     geom_hline(yintercept = 0, color = "gray70") +
     geom_vline(xintercept = 0, color = "gray70") +
     geom_point(aes(color = obesidad), alpha = 0.55, size = 3) +     scale_color_manual(values=c("black","red"))+ggtitle("Factor 1 Males BMI")+coord_cartesian(xlim = c(min(datagg[,1]),max(datagg[,1]))) 
  

  

ggarrange(p1,p2,p3,ncol=2,nrow=2)
```


```{r}
factores.all <- get_factors(model,factors = 1,scale = T)
p1 <- plot.methods3(factores.all$Female[,1],datos.common[datos.common$GROUP=="Female",])+ggtitle("Female")
p2 <- plot.methods3(factores.all$Male[,1],datos.common[datos.common$GROUP=="Male",])+ggtitle("Male")
p3 <- plot.methods3(factores.all$PCOS[,1],datos.common[datos.common$GROUP=="PCOS",])+ggtitle("PCOS")

ggarrange(p1,p2,p3,nrow=2,ncol=2)
```

### Correlacion covariables MUJERES
```{r}
correlate_factors_with_covariates(model, 
  covariates = c("obesidad","edad","BMI","WC","WHR","SHGB","hsCRP","TE2","E2"), 
  plot="log_pval",groups = "Female")
```


```{r}
datagg <- as.data.frame(cbind(edad=model@samples_metadata$edad[model@samples_metadata$grupo=="Female"],Factor1=get_factors(model,factors=1,scale = T,as.data.frame = T,groups = "Female")$value))

  obesidad<-as.factor(model@samples_metadata$obesidad[model@samples_metadata$grupo=="Female"])
  p1<-ggplot(datagg,aes(x=edad, y=Factor1)) +
     geom_hline(yintercept = 0, color = "gray70") +
     geom_vline(xintercept = 0, color = "gray70") +
     geom_point(aes(color = obesidad), alpha = 0.55, size = 3) +     scale_color_manual(values=c("black","red"))+ggtitle("Factor 1 Female edad")+coord_cartesian(xlim = c(min(datagg[,1]),max(datagg[,1]))) 
  
 datagg <- as.data.frame(cbind(WC=model@samples_metadata$WC[model@samples_metadata$grupo=="Female"],Factor1=get_factors(model,factors=1,scale = T,as.data.frame = T,groups = "Female")$value))

  obesidad<-as.factor(model@samples_metadata$obesidad[model@samples_metadata$grupo=="Female"])
  p2<-ggplot(datagg,aes(x=WC, y=Factor1)) +
     geom_hline(yintercept = 0, color = "gray70") +
     geom_vline(xintercept = 0, color = "gray70") +
     geom_point(aes(color = obesidad), alpha = 0.55, size = 3) +     scale_color_manual(values=c("black","red"))+ggtitle("Factor 1 Female WC")+coord_cartesian(xlim = c(min(datagg[,1]),max(datagg[,1]))) 
  
 datagg <- as.data.frame(cbind(BMI=model@samples_metadata$BMI[model@samples_metadata$grupo=="Female"],Factor1=get_factors(model,factors=1,scale = T,as.data.frame = T,groups = "Female")$value))

  obesidad<-as.factor(model@samples_metadata$obesidad[model@samples_metadata$grupo=="Female"])
  p3<-ggplot(datagg,aes(x=BMI, y=Factor1)) +
     geom_hline(yintercept = 0, color = "gray70") +
     geom_vline(xintercept = 0, color = "gray70") +
     geom_point(aes(color = obesidad), alpha = 0.55, size = 3) +     scale_color_manual(values=c("black","red"))+ggtitle("Factor 1 Female BMI")+coord_cartesian(xlim = c(min(datagg[,1]),max(datagg[,1]))) 
  
  datagg <- as.data.frame(cbind(E2=model@samples_metadata$E2[model@samples_metadata$grupo=="Female"],Factor1=get_factors(model,factors=1,scale = T,as.data.frame = T,groups = "Female")$value))

  obesidad<-as.factor(model@samples_metadata$obesidad[model@samples_metadata$grupo=="Female"])
  p4<-ggplot(datagg,aes(x=E2, y=Factor1)) +
     geom_hline(yintercept = 0, color = "gray70") +
     geom_vline(xintercept = 0, color = "gray70") +
     geom_point(aes(color = obesidad), alpha = 0.55, size = 3) +     scale_color_manual(values=c("black","red"))+ggtitle("Factor 1 Female E2")+coord_cartesian(xlim = c(min(datagg[,1]),max(datagg[,1]))) 
  
  

ggarrange(p1,p2,p3,p4,ncol=2,nrow=2)
```




### Correlacion covariables  pcos
```{r}
correlate_factors_with_covariates(model, 
  covariates = c("obesidad","edad","BMI","WC","WHR","SHGB","hsCRP","TE2","E2"), 
  plot="log_pval",groups = "PCOS")
```

```{r}
datagg <- as.data.frame(cbind(E2=model@samples_metadata$E2[model@samples_metadata$grupo=="PCOS"],Factor1=get_factors(model,factors=1,scale = T,as.data.frame = T,groups = "PCOS")$value))

  obesidad<-as.factor(model@samples_metadata$obesidad[model@samples_metadata$grupo=="PCOS"])
  p1<-ggplot(datagg,aes(x=E2, y=Factor1)) +
     geom_hline(yintercept = 0, color = "gray70") +
     geom_vline(xintercept = 0, color = "gray70") +
     geom_point(aes(color = obesidad), alpha = 0.55, size = 3) +     scale_color_manual(values=c("black","red"))+ggtitle("Factor 1 PCOS E2")+coord_cartesian(xlim = c(min(datagg[,1]),max(datagg[,1]))) 
  
 datagg <- as.data.frame(cbind(WC=model@samples_metadata$WC[model@samples_metadata$grupo=="PCOS"],Factor1=get_factors(model,factors=1,scale = T,as.data.frame = T,groups = "PCOS")$value))

  obesidad<-as.factor(model@samples_metadata$obesidad[model@samples_metadata$grupo=="PCOS"])
  p2<-ggplot(datagg,aes(x=WC, y=Factor1)) +
     geom_hline(yintercept = 0, color = "gray70") +
     geom_vline(xintercept = 0, color = "gray70") +
     geom_point(aes(color = obesidad), alpha = 0.55, size = 3) +     scale_color_manual(values=c("black","red"))+ggtitle("Factor 1 PCOS WC")+coord_cartesian(xlim = c(min(datagg[,1]),max(datagg[,1]))) 
  
 datagg <- as.data.frame(cbind(BMI=model@samples_metadata$BMI[model@samples_metadata$grupo=="PCOS"],Factor1=get_factors(model,factors=1,scale = T,as.data.frame = T,groups = "PCOS")$value))

  obesidad<-as.factor(model@samples_metadata$obesidad[model@samples_metadata$grupo=="PCOS"])
  p3<-ggplot(datagg,aes(x=BMI, y=Factor1)) +
     geom_hline(yintercept = 0, color = "gray70") +
     geom_vline(xintercept = 0, color = "gray70") +
     geom_point(aes(color = obesidad), alpha = 0.55, size = 3) +     scale_color_manual(values=c("black","red"))+ggtitle("Factor 1 PCOS BMI")+coord_cartesian(xlim = c(min(datagg[,1]),max(datagg[,1]))) 
  

   datagg <- as.data.frame(cbind(hsCRP=model@samples_metadata$hsCRP[model@samples_metadata$grupo=="PCOS"],Factor1=get_factors(model,factors=1,scale = T,as.data.frame = T,groups = "PCOS")$value))

  obesidad<-as.factor(model@samples_metadata$obesidad[model@samples_metadata$grupo=="PCOS"])
  p4<-ggplot(datagg,aes(x=hsCRP, y=Factor1)) +
     geom_hline(yintercept = 0, color = "gray70") +
     geom_vline(xintercept = 0, color = "gray70") +
     geom_point(aes(color = obesidad), alpha = 0.55, size = 3) +     scale_color_manual(values=c("black","red"))+ggtitle("Factor 1 PCOS hsCRP")+coord_cartesian(xlim = c(min(datagg[,1]),max(datagg[,1]))) 
  
ggarrange(p1,p2,p3,p4,ncol=2,nrow=2)
```

### Metaboloma



```{r}
top_weights <- plot_top_weights(model,scale = T,view =1)
top_weights<-data.frame(features=as.character(top_weights$data$feature),values=top_weights$data$value)

ggarrange(plot_top_weights(model,scale=T,view=1),plot_weights(model,scale=T,view=1,factor=1))
```

#### hOMBRES
```{r}
top_weights <- plot_top_weights(model,scale = T,view =1,factors = 1)
top_weights<-data.frame(features=as.character(top_weights$data$feature),values=top_weights$data$value)
plot_data_scatter(model,factor=1,view=1,features = top_weights$features,color_by ="obesidad",groups = "Male")


```


#### MUJERES

```{r}
plot_data_scatter(model,factor=1,view=1,features = top_weights$features,color_by ="obesidad",groups = "Female")


```





#### PCOS

```{r}
plot_data_scatter(model,factor=1,view=1,features = top_weights$features,color_by ="obesidad",groups = "PCOS")


```








### MetaGENOMA



```{r}
top_weights <- plot_top_weights(model,scale = T,view =2,factors =4 )
top_weights<-data.frame(features=as.character(top_weights$data$feature),values=top_weights$data$value)

ggarrange(plot_top_weights(model,scale=T,view=2,factors = 4),plot_weights(model,scale=T,view=,factor=4))
```

#### hOMBRES
```{r}
top_weights <- plot_top_weights(model,scale = T,view =2,factors = 4)
top_weights<-data.frame(features=as.character(top_weights$data$feature),values=top_weights$data$value)
plot_data_scatter(model,factor=4,view=2,features = top_weights$features,color_by ="obesidad",groups = "Male")


```


#### MUJERES

```{r}
top_weights <- plot_top_weights(model,scale = T,view =2,factors = 1)
top_weights<-data.frame(features=as.character(top_weights$data$feature),values=top_weights$data$value)
plot_data_scatter(model,factor=1,view=2,features = top_weights$features,color_by ="obesidad",groups = "Female")


```


```{r}
plot_data_scatter()
```



#### PCOS

```{r}
plot_data_scatter(model,factor=4,view=2,features = top_weights$features,color_by ="obesidad",groups = "Female")


```








## Mantel
```{r}
fn.explore_grupo<- function(datos.in){
datos.in<-t(datos_in)
datos <- as.data.frame(datos.in)
datos$obesidad <- variables_in_bacteria[variables_in_bacteria$Paciente %in% rownames(datos) ,"OBESE"]
descarte <--c(dim(datos)[2]:(dim(datos)[2]+1))
#### gruposde interes
obese <- datos[datos$obesidad=="Obese",descarte]
no.obese <- datos[datos$obesidad=="No Obese",descarte]
out <- list(obese=obese,
            no.obese=no.obese)
return(out)
}

```

### PCOS




```{r}
pesos.metaboloma <- unique(as.character(plot_top_weights(model,view = 1,factors = 1,scale = T)$data$feature))
pesos.metagenoma <- unique(as.character(plot_top_weights(model,view=2,factors = 4,scale = T)$data$feature))
pesos <- c(pesos.metaboloma,pesos.metagenoma)
datos_in <- rbind(scale(model@data$metaboloma$PCOS),scale(model@data$microbioma$PCOS))
aux <- fn.explore_grupo(t(datos_in))
names(aux)
library(ape)
combinaciones <- t(combn(names(aux),2))
p<-matrix(NA,nrow=10,ncol=1)
for(it in 1:10){
  

      
      p[it,ci] <- mantel.test(cor(aux[[combinaciones[1,1]]],method = "spearman"),
                           cor(aux[[combinaciones[1,2]]],method = "spearman"))$p
    
}

p<-colMeans(p)
names(p) <- apply(combinaciones, 1, function(x) paste(x[1],x[2]))
mantel.df <- data.frame(p=p.adjust(p,"bonferroni"),nombres=names(p))
mantel.df$asociacion <- ifelse(mantel.df$p<0.05,"<0.05",NA)
mantel.df$no_asociacion <-  ifelse(mantel.df$p>0.05,">0.05",NA)
# write.csv(mantel.df,"./mantel_interacciones.csv")
mantel.df

    # wi.p.matrix[, mc.i] <- wilcox.fast(t.input, setAs
```

```{r}
factors <- get_factors(model, as.data.frame = T,scale = F,groups = "PCOS",factors ="all")

weights <-  get_weights(model,as.data.frame=T,scale=F,views="all",factors = "all")
                 
factores.recast <-as.data.frame(reshape2::recast(data=factors,sample~factor))
rownames(factores.recast) <- factores.recast$sample
factores.recast <- factores.recast[,-1]
weights.recast <-as.data.frame(reshape2::recast(data=weights,feature~factor))
rownames(weights.recast) <- weights.recast$feature
weights.recast <- weights.recast[,-1]
weights.recast$views <- c(rep("metabooloma",37),rep("microbioma",94))

top_weights.metagenoma <- unique(as.character(plot_top_weights(model,scale = T,view =1,factors = "all")$data$feature))
top_weights.metaboloma <- unique(as.character(plot_top_weights(model,scale = T,view =2,factors = "all")$data$feature))

pesos <- c(top_weights.metagenoma,top_weights.metaboloma)

Y <-(as.matrix(factores.recast) %*% t(weights.recast[,-ncol(weights.recast)]))
aux <- fn.explore_grupo(Y)
library(ape)
      
     ( p <-mantel.test(cor(aux[[1]],method = "spearman"),
                           cor(aux[[2]],method = "spearman"))$p)
    
# names(p) <- apply(combinaciones, 1, function(x) paste(x[1],x[2]))


```



### MUJERES




```{r}
pesos.metaboloma <- unique(as.character(plot_top_weights(model,view = 1,factors = 1,scale = T)$data$feature))
pesos.metagenoma <- unique(as.character(plot_top_weights(model,view=2,factors = 4,scale = T)$data$feature))
pesos <- c(pesos.metaboloma,pesos.metagenoma)
datos_in <- rbind(scale(model@data$metaboloma$Female),scale(model@data$microbioma$Female))
aux <- fn.explore_grupo(t(datos_in)[,pesos])
names(aux)
library(ape)
combinaciones <- t(combn(names(aux),2))
p<-matrix(NA,nrow=10,ncol=1)
for(it in 1:10){
  

      
      p[it,ci] <- mantel.test(cor(aux[[combinaciones[1,1]]],method = "spearman"),
                           cor(aux[[combinaciones[1,2]]],method = "spearman"))$p
    
}

p<-colMeans(p)
names(p) <- apply(combinaciones, 1, function(x) paste(x[1],x[2]))
mantel.df <- data.frame(p=p.adjust(p,"bonferroni"),nombres=names(p))
mantel.df$asociacion <- ifelse(mantel.df$p<0.05,"<0.05",NA)
mantel.df$no_asociacion <-  ifelse(mantel.df$p>0.05,">0.05",NA)
# write.csv(mantel.df,"./mantel_interacciones.csv")
mantel.df

    # wi.p.matrix[, mc.i] <- wilcox.fast(t.input, setAs
```

```{r}
factors <- get_factors(model, as.data.frame = T,scale = F,groups = "Female",factors ="all")

weights <-  get_weights(model,as.data.frame=T,scale=F,views="all",factors = "all")
                 
factores.recast <-as.data.frame(reshape2::recast(data=factors,sample~factor))
rownames(factores.recast) <- factores.recast$sample
factores.recast <- factores.recast[,-1]
weights.recast <-as.data.frame(reshape2::recast(data=weights,feature~factor))
rownames(weights.recast) <- weights.recast$feature
weights.recast <- weights.recast[,-1]
weights.recast$views <- c(rep("metabooloma",37),rep("microbioma",94))

top_weights.metagenoma <- unique(as.character(plot_top_weights(model,scale = T,view =1,factors = "all")$data$feature))
top_weights.metaboloma <- unique(as.character(plot_top_weights(model,scale = T,view =2,factors = "all")$data$feature))

pesos <- c(top_weights.metagenoma,top_weights.metaboloma)

Y <-(as.matrix(factores.recast) %*% t(weights.recast[,-ncol(weights.recast)]))
aux <- fn.explore_grupo(Y)
library(ape)
      
     ( p <-mantel.test(cor(aux[[1]],method = "spearman"),
                           cor(aux[[2]],method = "spearman"))$p)
    
# names(p) <- apply(combinaciones, 1, function(x) paste(x[1],x[2]))


```












### HOMBRES




```{r}
pesos.metaboloma <- unique(as.character(plot_top_weights(model,view = 1,factors = 1,scale = T)$data$feature))
pesos.metagenoma <- unique(as.character(plot_top_weights(model,view=2,factors = 4,scale = T)$data$feature))
pesos <- c(pesos.metaboloma,pesos.metagenoma)
datos_in <- rbind(scale(model@data$metaboloma$Male),scale(model@data$microbioma$Male))
aux <- fn.explore_grupo(t(datos_in)[,pesos])
names(aux)
library(ape)
combinaciones <- t(combn(names(aux),2))
p<-matrix(NA,nrow=10,ncol=1)
for(it in 1:10){
  

      
      p[it,ci] <- mantel.test(cor(aux[[combinaciones[1,1]]],method = "spearman"),
                           cor(aux[[combinaciones[1,2]]],method = "spearman"))$p
    
}

p<-colMeans(p)
names(p) <- apply(combinaciones, 1, function(x) paste(x[1],x[2]))
mantel.df <- data.frame(p=p.adjust(p,"bonferroni"),nombres=names(p))
mantel.df$asociacion <- ifelse(mantel.df$p<0.05,"<0.05",NA)
mantel.df$no_asociacion <-  ifelse(mantel.df$p>0.05,">0.05",NA)
# write.csv(mantel.df,"./mantel_interacciones.csv")
mantel.df

    # wi.p.matrix[, mc.i] <- wilcox.fast(t.input, setAs
```

```{r}
factors <- get_factors(model, as.data.frame = T,scale = F,groups = "Male",factors ="all")

weights <-  get_weights(model,as.data.frame=T,scale=F,views="all",factors = "all")
                 
factores.recast <-as.data.frame(reshape2::recast(data=factors,sample~factor))
rownames(factores.recast) <- factores.recast$sample
factores.recast <- factores.recast[,-1]
weights.recast <-as.data.frame(reshape2::recast(data=weights,feature~factor))
rownames(weights.recast) <- weights.recast$feature
weights.recast <- weights.recast[,-1]
weights.recast$views <- c(rep("metabooloma",37),rep("microbioma",94))

top_weights.metagenoma <- unique(as.character(plot_top_weights(model,scale = T,view =1,factors = "all")$data$feature))
top_weights.metaboloma <- unique(as.character(plot_top_weights(model,scale = T,view =2,factors = "all")$data$feature))

pesos <- c(top_weights.metagenoma,top_weights.metaboloma)

Y <-(as.matrix(factores.recast) %*% t(weights.recast[,-ncol(weights.recast)]))
aux <- fn.explore_grupo(Y)
library(ape)
      
     ( p <-mantel.test(cor(aux[[1]],method = "spearman"),
                           cor(aux[[2]],method = "spearman"))$p)
    
# names(p) <- apply(combinaciones, 1, function(x) paste(x[1],x[2]))


```









